---
title: "Zoid - troubleshooting"
subtitle: "run date: `r format(Sys.time(), '%B %d, %Y')`"
author: "M Fisher"
date: 'written 2023-09-26'
output: 
  html_document:
    toc: yes
    toc_float: yes
---

## Goal:

Use zoid to understand how diet varies with instar size (measured as carapace width, mm). 
Diet information generated with DNA metabarcoding, values in the data matrix represent sequence abundance. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(vegan)
library(labdsv)
library(magrittr)
library(zoid)

# User inputs
resultsdir <- 'data/results'
```

```{r data, include=FALSE}
cw.df <- readRDS(here('data','metadata','carapace_widths_for_zoid.rds'))
```

Variation in carapace width looks like: 
```{r}
ggplot(cw.df, aes(x=CW_mm)) +
  geom_histogram() + theme_bw()
```


## All taxa, species or lowest tax level

```{r}
zoidIN.reads <- readRDS(here(resultsdir,'zoidIN_nReads_allTaxa_by_CW.rds'))
dim(zoidIN.reads)

```

```{r echo=FALSE}
zoidIN.reads %>%
  pivot_longer(1:(dim(zoidIN.reads)[2]-1),names_to="taxon",values_to="reads") %>%
  ggplot(aes(x=reads/100)) + geom_histogram() +
  theme_bw() + theme(axis.text.x=element_text(angle=45,hjust=1))
```

create the design matrix (site / site type information per observation, aka per crab) and the data matrix (read counts only)
```{r}
design_matrix = zoidIN.reads[,which(names(zoidIN.reads)=="CW_mm"), drop=FALSE]
data_matrix = zoidIN.reads[,which(names(zoidIN.reads)!="CW_mm")]
design_matrix$y = 1 # dummy variable
```

attempt to zoid!
```{r eval=FALSE}
fit_1_prey <- fit_zoid(formula = y ~ CW_mm, 
                      design_matrix = design_matrix, 
                      data_matrix = as.matrix(data_matrix),
                       overdispersion = TRUE,
                       chains=1,   # for testing; change to 4
                       iter=20)   # for testing; change to 5000
```


Chain 1: Rejecting initial value:
Chain 1:   Log probability evaluates to log(0), i.e. negative infinity.
Chain 1:   Stan can't start sampling from this initial value.
Chain 1: 
Chain 1: Initialization between (-2, 2) failed after 100 attempts. 
Chain 1:  Try specifying initial values, reducing ranges of constrained values, or reparameterizing the model.
[1] "Error : Initialization failed."
[1] "error occurred during calling the sampler; sampling not done"



## All taxa, family
```{r}
zoidIN.FAMreads <- readRDS(here(resultsdir,'zoidIN_nReads_allTaxaFamilies_by_CW.rds'))
dim(zoidIN.FAMreads)

```

```{r echo=FALSE}
zoidIN.FAMreads %>%
  pivot_longer(1:(dim(zoidIN.FAMreads)[2]-1),names_to="family",values_to="reads") %>%
  ggplot(aes(x=family,y=reads/1000)) + geom_boxplot() +
  theme_bw() + theme(axis.text.x=element_text(angle=45,hjust=1))
```

create the design matrix (site / site type information per observation, aka per crab) and the data matrix (read counts only)
```{r}
design_matrix = zoidIN.FAMreads[,which(names(zoidIN.FAMreads)=="CW_mm"), drop=FALSE]
data_matrix = zoidIN.FAMreads[,which(names(zoidIN.FAMreads)!="CW_mm")]
design_matrix$y = 1 # dummy variable
```


attempt to zoid!
```{r eval=FALSE}
fit_2_prey <- fit_zoid(formula = y ~ CW_mm, 
                      design_matrix = design_matrix, 
                      data_matrix = as.matrix(data_matrix),
                       overdispersion = TRUE,
                       chains=1,   # for testing; change to 4
                       iter=500)   # for testing; change to 5000
```

Chain 1: Rejecting initial value:
Chain 1:   Log probability evaluates to log(0), i.e. negative infinity.
Chain 1:   Stan can't start sampling from this initial value.
Chain 1: 
Chain 1: Initialization between (-2, 2) failed after 100 attempts. 
Chain 1:  Try specifying initial values, reducing ranges of constrained values, or reparameterizing the model.
[1] "Error : Initialization failed."
[1] "error occurred during calling the sampler; sampling not done"


## All taxa, class

```{r}
zoidIN.CLASSreads <- readRDS(here(resultsdir,'zoidIN_nReads_allTaxaCLASS_by_CW.rds'))
dim(zoidIN.CLASSreads)
```
```{r echo=FALSE}
zoidIN.CLASSreads %>%
  pivot_longer(1:(dim(zoidIN.CLASSreads)[2]-1),names_to="class",values_to="reads") %>%
  ggplot(aes(x=class,y=reads/1000)) + geom_boxplot() +
  theme_bw() + theme(axis.text.x=element_text(angle=45,hjust=1))
```

create the design matrix (site / site type information per observation, aka per crab) and the data matrix (read counts only)
```{r}
design_matrix = zoidIN.CLASSreads[,which(names(zoidIN.CLASSreads)=="CW_mm"), drop=FALSE]
data_matrix = zoidIN.CLASSreads[,which(names(zoidIN.CLASSreads)!="CW_mm")]
design_matrix$y = 1 # dummy variable
```



attempt to zoid!
```{r eval=FALSE}
fit_3_prey <- fit_zoid(formula = y ~ CW_mm, 
                      design_matrix = design_matrix, 
                      data_matrix = as.matrix(data_matrix),
                       overdispersion = TRUE,
                       chains=1,   # for testing; change to 4
                       iter=500)   # for testing; change to 5000
```

