---
title: "Dungeness Diet Appendix"
author: "M Fisher"
date: "2022-12-03"
output: html_document
---

# Description

Summary document to produce all figures for the appendix. Currently, figures are:

1. Percent hashes and percent reads removed for (1) missing data, (2) predator DNA, (3) overall.
2. Variability between technical replicates, by crab
3. Alpha diversity x read depth per crab




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(vegan)
library(magrittr)

# User inputs
indir   <- 'data/blast'
run.nums <- c(1)
marker  <- 'lerayXT'
blastdir  <- 'data/blast'
```
<br>

# Data Filtering

## % Removed
```{r message=FALSE, dat_premoved}
## filtering summary dataframe - produced in script 5b ##
for(i in seq(1,length(run.nums))){
  if(i==1){
    summary_dat <- read_csv(here(blastdir, paste0(marker,"_","r",run.nums[i],"_sample_taxonomy_filtering_summary.csv")))
  } else{
    summary_dat <- bind_rows(summary_dat,
                               read_csv(here(blastdir, paste0(marker,"_","r",run.nums[i],"_sample_taxonomy_filtering_summary.csv"))))
  }
}
```
```{r message=FALSE}
crab.totals <-summary_dat %>%
  group_by(sample_id) %>% summarise(total_hashes=length(unique(Hash)), total_reads=sum(nReads))

## all retained ##
plotdat.sum <- summary_dat %>%
  mutate(retained=ifelse(reason=="retained",1,0)) %>%
  group_by(sample_id,retained) %>%
  summarise(nreads=sum(nReads),
            nhashes=sum(length(unique(Hash))), .groups="drop") %>%
  left_join(crab.totals,by=c("sample_id")) %>%
  mutate(phashes=1-(nhashes/total_hashes),
         preads=1-(nreads/total_reads)) %>%
  dplyr::select(sample_id,retained,phashes,preads) %>%
  pivot_longer(cols=c(phashes,preads),names_to="metric",values_to="proportion") %>%
  rename(removed=retained) %>%
  pivot_longer(cols=c(removed),names_to="filter_type",values_to="value")
               
## predator DNA ##
plotdat.reason <- summary_dat %>%
  mutate(predatorDNA=ifelse(reason=="predator DNA",1,0)) %>%
  group_by(sample_id,predatorDNA) %>%
  summarise(nreads=sum(nReads),
            nhashes=sum(length(unique(Hash))), .groups="drop") %>%
  left_join(crab.totals,by=c("sample_id")) %>%
  mutate(phashes=nhashes/total_hashes,
         preads=nreads/total_reads) %>%
  dplyr::select(sample_id,predatorDNA,phashes,preads) %>%
  pivot_longer(cols=c(phashes,preads),names_to="metric",values_to="proportion")%>%
  pivot_longer(cols=c(predatorDNA),names_to="filter_type",values_to="value")

plotdat <- bind_rows(plotdat.sum,plotdat.reason) %>%
  mutate(filter_type=ifelse(filter_type=="predatorDNA","Predator DNA","Overall Removed")) %>%
  mutate(metric=ifelse(metric=="phashes","Prop. Hashes","Prop. Reads"))
plotdat$filter_type <- factor(plotdat$filter_type, levels=c("Predator DNA","Overall Removed"))
```

First, by sample. 
```{r fig.height=6, fig.width=12}
## combined plot ##
p.filter <- ggplot(plotdat,aes(x=sample_id,y=proportion, fill=as.factor(value))) + geom_col() + facet_grid(cols=vars(metric),rows=vars(filter_type), drop=TRUE) +
  labs(x="Sample (Replicates)",y="Proportion of Hashes / Reads") +
  scale_fill_manual(values=c("darkseagreen4","dodgerblue4"), name="") +
  theme_classic() + theme(axis.text.x=element_blank())
p.filter
```
<br>

Save plot.
```{r echo=TRUE}
png(here('figs','SuppA_Fig1-filtering.png'), res=200,width=3000,height=1000)
p.filter
dev.off()
```
<br>

Then, as a histogram. 
```{r}
p.filter2 <- ggplot(plotdat %>% filter(value==1),aes(x=proportion,fill=filter_type)) +
  geom_histogram() +
  facet_grid(rows=vars(metric)) +
  labs(y="N. Samples (Replicates)",x="") +
  scale_fill_manual(values=c("brown4","grey60"), name="") +
  scale_y_continuous(expand=c(0,0)) +
  theme_classic()
```
<br>

Save plot.
```{r echo=TRUE}
png(here('figs','SuppA_Fig1-filtering-histogram.png'), res=200,width=1000,height=800)
p.filter2
dev.off()
```
<br>
<br>

# PCR Replicates
## Variability






