---
title: "troubleshoot x-l crabs"
author: "M Fisher"
date: "2023-09-27"
output: html_document
---


```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(janitor)
library(cowplot)
library(grid)

knitr::opts_chunk$set(echo = TRUE,
                      theme_set(theme_bw()))

plot.sim <- function(sim.out,nd){
  data.frame(day=seq(1:(nd+1)),
             Weight=sim.out$W,
             Carapace_Width=sim.out$CW) %>%
    pivot_longer(cols=c(2:3), names_to="Metric") %>%
    ggplot(aes(x=day,y=value)) +
    geom_path() +
    facet_wrap(~Metric, scales="free") + ylab("mm / grams")
}


dge_dir <- '../../../DungiGraciisBioenergetics'
nrg_dir <- '../../../NRG'
```
```{r Part0, echo=FALSE, warning=FALSE}
  # install.packages("devtools")
  #library("devtools");devtools::install_github("kholsman/NRG")

  #library("NRG")
  #setwd("/Users/kholsman/GitHub/NRG/NRG")

setwd('C:/Users/mfisher5/Documents/NRG')
source("R/make.R")


# source("R/G_fun.R")
# source("R/bioE.R")
# source("R/load_data.R")
# source("R/cuml_curv.R")
# source("R/ftc_fun.R")
# source("R/find_ftcpar.R")
# source("R/hello.R")
# source("R/nrg.R")
# source("R/plot_curve.R")
# source("R/plot_curve3.R")
# source("R/profileT.R")
# source("R/profileW.R")
# source("R/resp_fun.R")
# source("R/rmd2md.R")
# source("R/sim_W.R")
# source("R/simul_ftc.R")
# source("R/ultils.R")
# source("R/waste_fun.R")
source("R/subscripts/molty_IP.R")
source("R/subscripts/molty_MI.R")
source("R/subscripts/crab_afdw.R")
```
<br>



## Test sim

For now, use temperature loggers from the intertidal in Willapa Bay, WA. Provided by WSG Crab Team

```{r echo=FALSE}
load(here(nrg_dir,'data','WillapaBayTemperature.rds'))
Tdat_raw <- WillapaBayTemperature$WB.oysterville.2021

Tdat <- Tdat_raw %>%
  group_by(date) %>%
  summarise(TempC=mean(TempC))

summerTdat <- Tdat %>% filter(ymd(date) < ymd('2021-10-01') & ymd(date) > ymd('2021-07-01'))
```

starting and ending weight observations
```{r}
W_start <- 0.004518174
W_end <-  1.494803
```

for now, make the crab a cannibal
```{r}
dcrab_data$Eprey <- dcrab_data$Epred
```

here are the parameter values
```{r}
parms <- crabPARMS_USE[1,]
parms
```

```{r}
sim_dat <- dcrab_data
sim_dat$propDiet <- 1

## temperatures
sim_dat$TempC <- summerTdat$TempC     ## temperatures

## starting crab weight
sim_dat$W     <- W_start         ## set W data for the simulation to the first observed weight


##
nd     <-  dim(summerTdat)[1] # days of simulation

## weights at start and end of simulation
Wobs   <- data.frame(
  day = c(1,nd),
  W   = c(W_start,W_end)) 
Wobs$date <- summerTdat$date[Wobs$day]

# simulation data frames
DCRBdat4sim   <- list(
                  Wtarget  = Wobs,
                  parSIM  = parms,
                  simTdat = summerTdat,
                  dataSIM  = sim_dat,
                  dataMOLT = dataMOLT$Cmagister)

sim_out       <- summerTdat %>% mutate(W=0)
```

simulate!
```{r}
test.sim <- sim_W(par=parms$RFR, dataIN=DCRBdat4sim, LL=F)
```
```{r echo=FALSE}
plot.sim(sim.out =test.sim,nd=nd)
```


Weight is... large.

```{r}
data.frame(day=rep(1:(nd+1), 2),
           type=c(rep("simulated",nd+1),rep("calc from CW",nd+1)),
           gAFDW=c(test.sim$W,sapply(test.sim$CW,cw2afdw))) %>%
  ggplot(aes(x=day,y=gAFDW)) +
  geom_point() + geom_path() +
  facet_wrap(~type, scales="free_y") + labs("RFR=1; pIloss=0; pIndigest=0")
```




### RFR to 0.75
```{r}
test.sim2 <- sim_W(par=0.75, dataIN=DCRBdat4sim, LL=F)
plot.sim(sim.out =test.sim2,nd=nd)
```
```{r}
data.frame(day=rep(1:(nd+1), 2),
           type=c(rep("simulated",nd+1),rep("calc from CW",nd+1)),
           gAFDW=c(test.sim2$W,sapply(test.sim2$CW,cw2afdw))) %>%
  ggplot(aes(x=day,y=gAFDW)) +
  geom_point() + geom_path() +
  facet_wrap(~type, scales="free_y") + labs(title="RFR=0.75")
```




### RFR to 0.75, I loss > 0

I found some graphs from Sean that make it look like Ingestion loss for Dungeness crab may vary across temperatures, but it mostly seems to be < 0.10
```{r}
sim_dat <- dcrab_data
sim_dat$propDiet <- 1
sim_dat$propIloss <- 0.10
```

```{r}
## temperatures
sim_dat$TempC <- summerTdat$TempC     ## temperatures

## starting crab weight
sim_dat$W     <- W_start         ## set W data for the simulation to the first observed weight


##
nd     <-  dim(summerTdat)[1] # days of simulation

## weights at start and end of simulation
Wobs   <- data.frame(
  day = c(1,nd),
  W   = c(W_start,W_end)) 
Wobs$date <- summerTdat$date[Wobs$day]

# simulation data frames
DCRBdat4sim   <- list(
                  Wtarget  = Wobs,
                  parSIM  = parms,
                  simTdat = summerTdat,
                  dataSIM  = sim_dat,
                  dataMOLT = dataMOLT$Cmagister)

sim_out       <- summerTdat %>% mutate(W=0)
```


```{r}
test.sim3 <- sim_W(par=0.75, dataIN=DCRBdat4sim, LL=F)
plot.sim(sim.out =test.sim3,nd=nd)
```


compare simulated weight to weight based on carapace width:
```{r}
data.frame(day=rep(1:(nd+1), 2),
           type=c(rep("simulated",nd+1),rep("calc from CW",nd+1)),
           gAFDW=c(test.sim3$W,sapply(test.sim3$CW,cw2afdw))) %>%
  ggplot(aes(x=day,y=gAFDW)) +
  geom_point() + geom_path() +
  facet_wrap(~type, scales="free_y") + labs(title="RFR=0.75; prop Ingest loss=0.10")
```

### RFR to 0.4, I loss > 0, Indigest > 0

Just making up indigestable biomass of consumed prey as 0.40
```{r}
sim_dat <- dcrab_data
sim_dat$propDiet <- 1
sim_dat$propIloss <- 0.10
sim_dat$propIndigest <- 0.40
```

```{r}
## temperatures
sim_dat$TempC <- summerTdat$TempC     ## temperatures

## starting crab weight
sim_dat$W     <- W_start         ## set W data for the simulation to the first observed weight


##
nd     <-  dim(summerTdat)[1] # days of simulation

## weights at start and end of simulation
Wobs   <- data.frame(
  day = c(1,nd),
  W   = c(W_start,W_end)) 
Wobs$date <- summerTdat$date[Wobs$day]

# simulation data frames
DCRBdat4sim   <- list(
                  Wtarget  = Wobs,
                  parSIM  = parms,
                  simTdat = summerTdat,
                  dataSIM  = sim_dat,
                  dataMOLT = dataMOLT$Cmagister)

sim_out       <- summerTdat %>% mutate(W=0)
```


```{r}
test.sim4 <- sim_W(par=0.4, dataIN=DCRBdat4sim, LL=F)
plot.sim(sim.out =test.sim4,nd=nd)
```


compare simulated weight to weight based on carapace width:
```{r}
data.frame(day=rep(1:(nd+1), 2),
           type=c(rep("simulated",nd+1),rep("calc from CW",nd+1)),
           gAFDW=c(test.sim4$W,sapply(test.sim4$CW,cw2afdw))) %>%
  ggplot(aes(x=day,y=gAFDW)) +
  geom_point() + geom_path() +
  facet_wrap(~type, scales="free_y") + labs("RFR=0.75; pIloss=0.10; pIndigest=0.40")
```


Nothing seems to be changing for proportion of indigestible prey > 0? But regardless, g AFDW for the simulation is too large.




## bioE

Plot growth / consumption / respiration across temperature range
```{r}
tmp.g <- data.frame(bioE(par=parms,
                                  data=dcrab_data)[[2]]) %>%
  dplyr::select(TempC,G_ggd,G_som_ggd,G_lipid_ggd,G_shell_ggd,
                Cmax_ggd,C_ggd,
                R_ggd,R_Act_ggd)

tmp.g.long <- tmp.g %>% pivot_longer(cols=c(2:ncol(tmp.g)), names_to="variable", values_to="ggd")
```
```{r echo=FALSE}
tmp.g.long %>% filter(variable %in% c("G_ggd","G_som_ggd","G_lipid_ggd","G_shell_ggd")) %>%
ggplot(aes(x=TempC,y=ggd,col=variable)) +
  geom_point(size=0.5) +
  geom_path() +
  geom_hline(aes(yintercept=0)) +
  labs(subtitle="RFR=1, Ingestion loss 0")

tmp.g.long %>% filter(variable %in% c("Cmax_ggd","C_ggd","R_ggd","R_Act_ggd")) %>%
ggplot(aes(x=TempC,y=ggd,col=variable)) +
  geom_point(size=0.5) +
  geom_path() +
  geom_hline(aes(yintercept=0)) +
  labs(subtitle="RFR=1, Ingestion loss 0")
```

As in the last simulation, reduce RFR to 0.4, make Ingestion loss 10%

```{r}
sim_dat <- dcrab_data
sim_dat$propDiet <- 1
sim_dat$propIloss <- 0.10

parms2 <- parms
parms2$RFR <- 0.40


## rerun bioE
tmp.g2 <- data.frame(bioE(par=parms2,
                                  data=sim_dat)[[2]]) %>%
  dplyr::select(TempC,G_ggd,G_som_ggd,G_lipid_ggd,G_shell_ggd,
                Cmax_ggd,C_ggd,
                R_ggd,R_Act_ggd)

tmp.g2.long <- tmp.g2 %>% pivot_longer(cols=c(2:ncol(tmp.g2)), names_to="variable", values_to="ggd")
```

```{r echo=FALSE}
g.subtitle <- "RFR=0.40, Ingestion loss 0.10"
tmp.g2.long %>% filter(variable %in% c("G_ggd","G_som_ggd","G_lipid_ggd","G_shell_ggd")) %>%
ggplot(aes(x=TempC,y=ggd,col=variable)) +
  geom_point(size=0.5) +
  geom_path() +
  geom_hline(aes(yintercept=0)) +
  labs(subtitle=g.subtitle)

tmp.g2.long %>% filter(variable %in% c("Cmax_ggd","C_ggd","R_ggd","R_Act_ggd")) %>%
ggplot(aes(x=TempC,y=ggd,col=variable)) +
  geom_point(size=0.5) +
  geom_path() +
  geom_hline(aes(yintercept=0)) +
  labs(subtitle=g.subtitle)
```

Consumption is waaaay higher than respiration. 

There is a note where I was supposed to make sure the Act intercept parameter in the respiration function (`11.336`) was in the correct units. It looks like I had to convert respiration from g AFDW / hr to g AFDW / day. Maybe I needed to also do this with the `Act` intercept? Test this by adjusting the `customresp` aregument in the crab dataframe.

```{r}
sim_dat <- dcrab_data

sim_dat$customresp <- function(par,data){
  lnR_ulO2_gAFDWhr  <- par$RA*(data$TempC^2) + (par$RB*data$TempC*log(data$W + 10^-9)) + (data$RC*data$TempC) + data$RD
  R_ulO2_gAFDWhr    <- exp(lnR_ulO2_gAFDWhr)
  R_ulO2_gAFDWhr *24  -> R_ulO2_gAFDWd
  R_ulO2_gAFDWd*0.001 -> R_mlO2_gAFDWd
  R_mlO2_gAFDWd*4.8   -> R_cal_gAFDWd
  R_cal_gAFDWd*4.186  -> R_jgAFDWd
  R_jgAFDWd/1000      -> R_kjgAFDWd
  Act <- (0.147*data$W) + (-0.533*data$TempC) + (11.336*24)
  # return(list(R_jgd=R_kjgAFDWd, Act=Act))
  return(list(R_jgd=R_jgAFDWd, Act=Act))
}


## rerun bioE
tmp.g3 <- data.frame(bioE(par=parms,
                                  data=sim_dat)[[2]]) %>%
  dplyr::select(TempC,G_ggd,G_som_ggd,G_lipid_ggd,G_shell_ggd,
                Cmax_ggd,C_ggd,
                R_ggd,R_Act_ggd)

tmp.g3.long <- tmp.g3 %>% pivot_longer(cols=c(2:ncol(tmp.g3)), names_to="variable", values_to="ggd")
```

```{r echo=FALSE}
tmp.g3.long %>% filter(variable %in% c("G_ggd","G_som_ggd","G_lipid_ggd","G_shell_ggd")) %>%
ggplot(aes(x=TempC,y=ggd,col=variable)) +
  geom_point(size=0.5) +
  geom_path() +
  geom_hline(aes(yintercept=0)) +
  labs(subtitle="Act intercept * 24hr")

tmp.g3.long %>% filter(variable %in% c("Cmax_ggd","C_ggd","R_ggd","R_Act_ggd")) %>%
ggplot(aes(x=TempC,y=ggd,col=variable)) +
  geom_point(size=0.5) +
  geom_path() +
  geom_hline(aes(yintercept=0)) +
  labs(subtitle="Act intercept * 24hr")
```


And when RFR / Prop I Loss are lower?
```{r}
sim_dat <- dcrab_data
sim_dat$propDiet <- 1
sim_dat$propIloss <- 0.10
sim_dat$propIndigest <- 0.40

# parms2 <- parms
# parms2$RFR <- 0.40

sim_dat$customresp <- function(par,data){
  lnR_ulO2_gAFDWhr  <- par$RA*(data$TempC^2) + (par$RB*data$TempC*log(data$W + 10^-9)) + (data$RC*data$TempC) + data$RD
  R_ulO2_gAFDWhr    <- exp(lnR_ulO2_gAFDWhr)
  R_ulO2_gAFDWhr *24  -> R_ulO2_gAFDWd
  R_ulO2_gAFDWd*0.001 -> R_mlO2_gAFDWd
  R_mlO2_gAFDWd*4.8   -> R_cal_gAFDWd
  R_cal_gAFDWd*4.186  -> R_jgAFDWd
  R_jgAFDWd/1000      -> R_kjgAFDWd
  Act <- (0.147*data$W) + (-0.533*data$TempC) + (11.336*24)
  # return(list(R_jgd=R_kjgAFDWd, Act=Act))
  return(list(R_jgd=R_jgAFDWd, Act=Act))
}

## rerun bioE
tmp.g4 <- data.frame(bioE(par=parms2,
                                  data=sim_dat)[[2]]) %>%
  dplyr::select(TempC,G_ggd,G_som_ggd,G_lipid_ggd,G_shell_ggd,
                Cmax_ggd,C_ggd,
                R_ggd,R_Act_ggd)

tmp.g4.long <- tmp.g4 %>% pivot_longer(cols=c(2:ncol(tmp.g4)), names_to="variable", values_to="ggd")
```

```{r echo=FALSE}
g.subtitle <- "Act Intercept Ajudst; RFR=0.40, Ingestion loss 0.10"
tmp.g4.long %>% filter(variable %in% c("G_ggd","G_som_ggd","G_lipid_ggd","G_shell_ggd")) %>%
ggplot(aes(x=TempC,y=ggd,col=variable)) +
  geom_point(size=0.5) +
  geom_path() +
  geom_hline(aes(yintercept=0)) +
  labs(subtitle=g.subtitle)

tmp.g4.long %>% filter(variable %in% c("Cmax_ggd","C_ggd","R_ggd","R_Act_ggd")) %>%
ggplot(aes(x=TempC,y=ggd,col=variable)) +
  geom_point(size=0.5) +
  geom_path() +
  geom_hline(aes(yintercept=0)) +
  labs(subtitle=g.subtitle)
```

Check that this doesn't make a difference in the simulation

```{r}
sim_dat <- dcrab_data
sim_dat$propDiet <- 1
sim_dat$propIloss <- 0.10
sim_dat$propIndigest <- 0.40

sim_dat$customresp <- function(par,data){
  lnR_ulO2_gAFDWhr  <- par$RA*(data$TempC^2) + (par$RB*data$TempC*log(data$W + 10^-9)) + (data$RC*data$TempC) + data$RD
  R_ulO2_gAFDWhr    <- exp(lnR_ulO2_gAFDWhr)
  R_ulO2_gAFDWhr *24  -> R_ulO2_gAFDWd
  R_ulO2_gAFDWd*0.001 -> R_mlO2_gAFDWd
  R_mlO2_gAFDWd*4.8   -> R_cal_gAFDWd
  R_cal_gAFDWd*4.186  -> R_jgAFDWd
  R_jgAFDWd/1000      -> R_kjgAFDWd
  Act <- (0.147*data$W) + (-0.533*data$TempC) + (11.336*24)
  # return(list(R_jgd=R_kjgAFDWd, Act=Act))
  return(list(R_jgd=R_jgAFDWd, Act=Act))
}
```

```{r}
## temperatures
sim_dat$TempC <- summerTdat$TempC     ## temperatures

## starting crab weight
sim_dat$W     <- W_start         ## set W data for the simulation to the first observed weight


##
nd     <-  dim(summerTdat)[1] # days of simulation

## weights at start and end of simulation
Wobs   <- data.frame(
  day = c(1,nd),
  W   = c(W_start,W_end)) 
Wobs$date <- summerTdat$date[Wobs$day]

# simulation data frames
DCRBdat4sim   <- list(
                  Wtarget  = Wobs,
                  parSIM  = parms,
                  simTdat = summerTdat,
                  dataSIM  = sim_dat,
                  dataMOLT = dataMOLT$Cmagister)

sim_out       <- summerTdat %>% mutate(W=0)
```


```{r}
test.sim4 <- sim_W(par=0.4, dataIN=DCRBdat4sim, LL=F)
plot.sim(sim.out =test.sim4,nd=nd)
```
