---
title: "Creating Model Diets"
author: "M Fisher"
date: "2023-09-05"
output: html_document
---

# Description

From the dDNA data, summarize model diets for two analyses: (1) potential site-specific growth differences associated with different diets, (2) potential growth differences associated with diet item OA sensitivity.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(vegan)
library(magrittr)

# User inputs
blastdir   <- 'data/blast'
bedir <- 'data/bioE'
run.nums <- c(1,2)
marker  <- 'lerayXT'
```

```{r data, include=FALSE}
## filtered taxa ##
dat <- read_csv(here(blastdir, paste0(marker,"_","r",paste(run.nums,collapse="-"),"_sample_taxonomy_filtered2_uniqueTaxa.csv")))

dat %<>%
  mutate(site=ifelse(site=="KAY" & crab_num > 18, "KAY-shell",site)) %>%
  mutate(estuary=ifelse(site %in% c("KAY","KAY-shell"),"Port Susan Bay",
                                    ifelse(site %in% c("LAR","SAMT","SAMI"), "Samish Bay","Padilla Bay")))

```
<br>

## Site Diets

For sites with > 10 crab, find the most commonly detected prey items (according to frequency of occurrence).



## OA sensitivity diets

Produce four diets: Hi/Lo Energy Density x Hi/Lo OA Sensitivity










