---
title: "Dungeness crab bioenergetics: updating model"
author: "Kirstin Holsman"
output:
  html_document:
    df_print: kable
    fig_caption: yes
    theme: flatly
    toc: true
    toc_float: true
  pdf_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 5
    highlight: tango
    keep_tex: yes
    latex_engine: xelatex
    toc: TRUE
  header-includes:
  - \usepackage{inputenc}
  - \usepackage{unicode-math}
  - \pagenumbering{gobble}
  word_document:
    fig_caption: yes
    fig_width: 4
    keep_md: yes
---


# Description

Make the following updates to the original Holsman et al. (2003) model: 

1. add ingestion (done)

2. new simulations for molting (done)

3. using the find_ftcpar function in Part 5 of the NRG vignette, re-fit consumption parameters Tco and Qc (in progress)

*units??*

4. 

```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(janitor)

knitr::opts_chunk$set(echo = TRUE,
                      theme_set(theme_bw()))

dge_dir <- '../../../DungiGraciisBioenergetics'
nrg_dir <- '../../../NRG'
```
```{r Part0, echo=FALSE, include=FALSE}
  # install.packages("devtools")
  #library("devtools");devtools::install_github("kholsman/NRG")

  #library("NRG")
  #setwd("/Users/kholsman/GitHub/NRG/NRG")

setwd('C:/Users/mfisher5/Documents/NRG')
source("R/make.R")


# source("R/G_fun.R")
# source("R/bioE.R")
# source("R/load_data.R")
# source("R/cuml_curv.R")
# source("R/ftc_fun.R")
# source("R/find_ftcpar.R")
# source("R/hello.R")
# source("R/nrg.R")
# source("R/plot_curve.R")
# source("R/plot_curve3.R")
# source("R/profileT.R")
# source("R/profileW.R")
# source("R/resp_fun.R")
# source("R/rmd2md.R")
# source("R/sim_W.R")
# source("R/simul_ftc.R")
# source("R/ultils.R")
# source("R/waste_fun.R")
source("R/subscripts/molty_IP.R")
source("R/subscripts/molty_MI.R")
source("R/subscripts/crab_afdw.R")
```
<br>


# Data

## consumption experiments

All of the experimental data from 2018 has been cleaned up from the original spreadsheet. The data frame contains three separate trials (and so tank numbers are repeated) 

```{r}
dat <- read_csv(here::here(dge_dir,"data","Experiments2018.csv")) %>% filter(species=="Dungeness")

str(dat)
```
<br>


For trial 1, 
Calculate the ingestion loss (I), egestion (F), consumption (C), and specific consumption (in gWW prey / gWW predator). 
```{r}
cdat <- dat %>%
  filter(trial==1) %>%
  mutate(I=i_filtered_w_nytex_clip-i_nytex_clip,
                F=f_filtered_w_nytex_clip-f_nytex_clip,
                C=total_squid_added - (squid_disc_left + I),  ## Note that in experimental data sheet, C calculation doesn't account for I
                C_gg=C/crab_mass_g)
head(dplyr::select(cdat, crab_mass_g, total_squid_added, squid_disc_left, I, F, C, C_gg))
```
<br>

### option 1: consumption in gPrey/gPred

Fit the new parameters using consumption from the experiments, as g Prey AFDW / g Pred AFDW. 


Re-calculate the C_gg value, so that it is in ash-free dry weight (like the model; currently in WW). I can first calculate crab AFDW using the carapace width for Dungeness crab (in Holsman et al. 2003). 
```{r}
cdat_AFDW <- cdat %>%
  mutate(crab_mass_gAFDW=(0.00005068 * (as.numeric(cw_inside_spines)^(2.79))))
```

I don't have a good conversion factor for squid, but I do have an average % water (of WW) and an average % Ash (of DW) across a range of squid species, from: Nancy Davis 1993. Caloric content of oceanic zooplankton and fishes for studies of salmonid food habits and their ecologically related species. (NPAFC Doe.) FRI-UW-93 12. Fisheries Research Institute, University of Washington, Seattle. 10 pp

Can I use these averages as conversion factors? g squid WW * (ww2dw as average % water) * (dw2afdw as average % ash)
```{r}
ww2dw <- mean(0.69,0.72,0.817,0.852)
dw2afdw <- mean(0.07,0.08,0.094,0.064,0.092)

cdat_AFDW %<>% mutate(C_gAFDW=(total_squid_added - squid_disc_left)*ww2dw*dw2afdw) %>%
  mutate(C_ggAFDW=C_gAFDW/crab_mass_gAFDW)

head(dplyr::select(cdat_AFDW, temp,crab_mass_g, crab_mass_gAFDW,C, C_gAFDW, C_gg, C_ggAFDW))
```


The value `C_ggAFDW` is what is put into the find_ftcpar function.



### option 2: Percent max consumption

Use the parameters from the model to calculate **theoretical** maximum possible consumption (will be in g AFDW) from crab mass and temperature. Then divide specific consumption by maximum consumption, to get proportion max consumption. Since it's a proportion, I don't have to convert units, I think?


Note that this does require using the estimated (not fitted) values for $Tco$ and $QC$, when calculating max consumption. 

in the `bioE` function for NRG, maximum specific consumption rate at temperature T = $C_{max}$ =  $C_A * W^{C_B} *f(T_c) * E_{prey}$. And $f(T_c)$ is created using the NRG function `fTC_fun`. This gives maximum consumption as J / g pred AFDW / day


First, get the values for $f(T_c)$. This requires the input parameters, and a data frame with the temperatures from the experiment.
```{r}
inparms <- crabPARMS_USE %>% filter(X=="Cmagister")

tdat <- data.frame(TempC=cdat$temp)
ftc_dat <- fTC_fun(par=inparms,data=tdat)

cdat %<>% bind_cols(ftc=ftc_dat)
```

Then, calculate "W" (g AFDW) from the carapace width, using equation 6 from Holsman et al. 
```{r}
cdat %<>% mutate(crab_mass_gAFDW=(0.00005068 * (as.numeric(cw_inside_spines)^(2.79))))
```

Now we have $f(T_c)$ and $W$, and we already have $C_A$ and $C_B$ in the parameters data frame. Putting that altogether we get $C_{max}$ in g Pred AFDW:
```{r}
cdat %<>% 
  mutate(Cmax=inparms$CA*(crab_mass_gAFDW^inparms$CB)*ftc)
head(dplyr::select(cdat, temp, crab_mass_g, C_gg, ftc,Cmax))
```

Convert $C_{max}$ to g Pred WW. 

**For this I needed a conversion factor** so I ended up using the conversion factor 0.123, which was calculated for *Carcinus maenas* by Gogina et al. (2022); and then I calculated proportion of max consumption by dividing C_max g PredWW by specific consumption (`C_gg`, which is in gPrey WW / g Pred WW).

Finally, multiply the proportion of max consumption by the max consumption value in g AFDW. 
```{r}
cdat %<>%
  mutate(Cmax_WW=Cmax*0.123) %>%
  mutate(Cp=C_gg/Cmax_WW) %>%
  mutate(C_ggAFDW=Cp*Cmax)

head(dplyr::select(cdat, temp, crab_mass_g, C,C_gg, ftc,Cmax,Cmax_WW,Cp,C_ggAFDW))
```

```{r echo=FALSE,fig.width=3, fig.height=3}
ggplot(cdat, aes(x=as.factor(temp),y=Cp)) + geom_violin() + geom_jitter(width=0.1)
```

Let's assume that the top point is an outlier, and remove it from the dataset that will be used to fit new parameters
```{r}
cdat_fitting <- filter(cdat, Cp < 1)
```

The value `C_ggAFDW` is what is put into the find_ftcpar function.




# Fit consumption

estimate the consumption curves  - and re-fit consumption parameters Tco and Qc - using MCMC to draw from informly distributed intial conditions.


### option 1: consumption in gPrey/gPred AFDW

```{r}
newDat1 <- cdat_AFDW %>%
  mutate(N_fed=115) %>%
  rename(Temp=temp,Size_CW=cw_widest,W_g=crab_mass_gAFDW,C_g=C_gAFDW,C_ggd=C_ggAFDW) %>%
  dplyr::select(Temp,Size_CW,W_g,C_g,C_ggd,I,F)
newDat1
```

make sure that the function `find_ftcpar` works with some changes to input names.
```{r}

inparms <- crabPARMS_USE %>% filter(X=="Cmagister")

indat <- list(
    logitlink    = TRUE,
    fitsigma     = FALSE,
    sigset       = 0.02,
    TempCIN      = newDat1$Temp,
    ftc_obsIN    = newDat1$C_ggd)
```

Is the function working?
```{r}
find_ftcpar(
  parms_use    = inparms,
  par = c(
    logTco   = log(8),
    logQc    = log(2),
    logsigma = log(.0002)),
  data=indat)
```

```{r}
m<-optim(fn=find_ftcpar,par=c(logTco=log(8),logQc=log(2),
                              logsigma=log(.0002)),parms_use = inparms,data=indat,
         hessian=TRUE,control=list(maxit=1e6))
# vc <- solve(m$hessian)
# se<-(sqrt(diag(vc)))


round(exp(m$par),3)
```

How close are these to our original values?
```{r}
inparms$Tco ; inparms$QC
```

yikes...


### option 2: consumption as proportion of maximum consumption
```{r}
newDat2 <- cdat_fitting %>%
  mutate(N_fed=115) %>%
  rename(Temp=temp,W_g=crab_mass_g,C_g=C_gg,C_ggd=C_ggAFDW) %>%
  dplyr::select(Temp,C_ggd,I,F)
newDat2
```

make sure that the function `find_ftcpar` works with some changes to input names.
```{r}
inparms <- crabPARMS_USE %>% filter(X=="Cmagister")

indat <- list(
    logitlink    = TRUE,
    fitsigma     = FALSE,
    sigset       = 0.02,
    TempCIN      = newDat2$Temp,
    ftc_obsIN    = newDat2$C_ggd)
```

Is the function working?
```{r}
find_ftcpar(
  parms_use    = inparms,
  par = c(
    logTco   = log(8),
    logQc    = log(2),
    logsigma = log(.0002)),
  data=indat)
```

```{r}
m2<-optim(fn=find_ftcpar,par=c(logTco=log(8),logQc=log(2),
                              logsigma=log(.0002)),parms_use = inparms,data=indat,
         hessian=TRUE,control=list(maxit=1e6))
# vc <- solve(m$hessian)
# se<-(sqrt(diag(vc)))


round(exp(m2$par),3)
```

How close are these to our original values?
```{r}
inparms$Tco ; inparms$QC
```

argh...


# Check CW --> W conversion

The model has a function to convert CW to W, and I want to check this against my measurements in the lab.
First, read in the lab data sheet and grab the CW (mm) and W (g wet weight) values
```{r}
sizedat <- read_csv(here('../../','data','bioE','instar_WW.csv'), skip=1) %>% dplyr::select(site, quadrat, CW_mm, crab_gWW)
```

Calculate the carapace widths according to the weights, and compare against measured carapace width. note that i have to use a second conversion equation to get from AFDW to WW
```{r}
sizedat %<>%
  mutate(crab_gAFDW_fromCW=cw2afdw_vectorize(v=CW_mm)) %>%
  mutate(crab_gWW_fromCW=afdw2ww_vectorize(v=crab_gAFDW_fromCW))
```
```{r echo=FALSE}
rlabel <- round(summary(lm(crab_gWW_fromCW ~ crab_gWW, data=sizedat))$adj.r.squared,2)
ggplot(sizedat, aes(x=crab_gWW, y=crab_gWW_fromCW)) + 
  geom_point() +
  geom_smooth(method="lm") + 
  geom_text(data=data.frame(x=0.2, y=0.5,label=paste0("r2=",rlabel)),aes(x=x,y=y,label=label))
```

Not bad! The scale I was working with wasn't the greatest. 



# Test run Holsman model



## temperature

For now, let's use intertidal temperature (from WSG crab team) from Oysterville in Willapa Bay
```{r}
load(here(nrg_dir,'data','WillapaBayTemperature.rds'))
str(WillapaBayTemperature)
```

Get a daily average temperature
```{r fig.width=4, fig.height=3}
Tdat_raw <- WillapaBayTemperature$WB.oysterville.2021
Tdat <- Tdat_raw %>%
  group_by(date) %>%
  summarise(TempC=mean(TempC))
ggplot() + geom_point(data=Tdat_raw, aes(x=date, y=TempC), col="grey60") +
  geom_path(data=Tdat, aes(x=date,y=TempC, group=1), col="darkred", linewidth=2)
```
<br>


## with original parameters

Dungeness crab parameters
```{r}
parms <- crabPARMS_USE %>% filter(X=="Cmagister"); str(parms)
```
molt information is in its own data frame:
```{r}
dataMOLT$Cmagister
```


Set up the simulation with some fake weights
```{r}
nd     <-  dim(Tdat)[1] # days of simulation
# at start of simulation
Wstart <- 2           

## made up simulation day, weights + date 
Wobs   <- data.frame(
  day = c(1,nd),
  W   = c(Wstart,16)) 
Wobs$date <- Tdat$date[Wobs$day]

# simulation data
sim_dat       <- dcrab_data
sim_dat$TempC <- Tdat$TempC     ## temperatures
sim_dat$W     <- Wstart         ## set W data for the simulation to the first observed weight

DCRBdat4sim   <- list(
                  Wtarget  = Wobs,
                  parSIM  = parms,
                  simTdat = Tdat,
                  dataSIM  = sim_dat,
                  dataMOLT = dataMOLT$Cmagister)

sim_out       <- Tdat %>% mutate(W=0)
```
```{r}
dcrb.oy.sim <- sim_W(par=0.4, dataIN=DCRBdat4sim, LL=F)
```

```{r echo=FALSE}
plotdat <- data.frame(day=seq(1:(nd+1)),
                      Weight=dcrb.oy.sim$W,
                      Carapace_Width=dcrb.oy.sim$CW) %>%
  pivot_longer(cols=c(2:3), names_to="Metric")
ggplot(plotdat,aes(x=day,y=value)) +
  geom_path() +
  facet_wrap(~Metric, scales="free") + ylab("mm / grams")
```

Weight seems high, will need to come back to that. 


## with newly fit parameters

Dungeness crab parameters
```{r eval=FALSE}
parms2 <- crabPARMS_USE %>% filter(X=="Cmagister")
parms2$Tco <- exp(m$par["Tco"])
  parms2$QC <- exp(m$par["QC"])
```
molt information is in its own data frame:
```{r eval=FALSE}
dataMOLT$Cmagister
```


Set up the simulation with some fake weights
```{r eval=FALSE}
nd     <-  dim(Tdat)[1] # days of simulation
# at start of simulation
Wstart <- 2           

## made up simulation day, weights + date 
Wobs   <- data.frame(
  day = c(1,nd),
  W   = c(Wstart,16)) 
Wobs$date <- Tdat$date[Wobs$day]

# simulation data
sim_dat       <- dcrab_data
sim_dat$TempC <- Tdat$TempC     ## temperatures
sim_dat$W     <- Wstart         ## set W data for the simulation to the first observed weight

DCRBdat4sim   <- list(
                  Wtarget  = Wobs,
                  parSIM  = parms2,
                  simTdat = Tdat,
                  dataSIM  = sim_dat,
                  dataMOLT = dataMOLT$Cmagister)

sim_out       <- Tdat %>% mutate(W=0)
```
```{r eval=FALSE}
dcrb.oy.sim <- sim_W(par=0.4, dataIN=DCRBdat4sim, LL=F)
```

```{r echo=FALSE, eval=FALSE}
plotdat <- data.frame(day=seq(1:(nd+1)),
                      Weight=dcrb.oy.sim$W,
                      Carapace_Width=dcrb.oy.sim$CW) %>%
  pivot_longer(cols=c(2:3), names_to="Metric")
ggplot(plotdat,aes(x=day,y=value)) +
  geom_path() +
  facet_wrap(~Metric, scales="free") + ylab("mm / grams")
```



# parking lot

Let's just take an average of the values for whole squid reported by: Nancy Davis 1993. Caloric content of oceanic zooplankton and fishes for studies of salmonid food habits and their ecologically related species. (NPAFC Doe.) FRI-UW-93 12. Fisheries Research Institute, University of Washington, Seattle. 10 pp
```{r}
squid.df <- data.frame(species=c("Illex illacabrosus","Illex illacabrosus","Illex illacabrosus","Loligo paelei","Loligo paelei","Loligo paelei","Loligo bravis","Squids"),
                           sample_area=c("NW atlantic","NW atlantic","NW atlantic","NW atlantic","NW atlantic","NW atlantic","NE Atlantic","Bristol Bay"),
                           calgWW=c(1695,NA,NA,1337,NA,NA,1051,NA),
                           calgAFDW=c(5636,5875,5397,5110,4991,5540,6342,5578)) %>%
  mutate(JgAFDW=calgAFDW*4.184)
squid_ed=mean(squid.df$JgAFDW); squid_ed
```