---
title: "Dungeness crab Bioenergetic - run OA diets"
author: "Kirstin Holsman, Mary Fisher"
output:
  html_document:
    df_print: kable
    fig_caption: yes
    theme: flatly
    toc: true
    toc_float: true
  pdf_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 5
    highlight: tango
    keep_tex: yes
    latex_engine: xelatex
    toc: TRUE
  header-includes:
  - \usepackage{inputenc}
  - \usepackage{unicode-math}
  - \pagenumbering{gobble}
  word_document:
    fig_caption: yes
    fig_width: 4
    keep_md: yes
---

# Description

Compare growth rates of crab with OA diets. 

To Do List:

- check weight calculation in sim_W function -- weight is too high!


```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(janitor)
library(cowplot)
library(grid)

knitr::opts_chunk$set(echo = TRUE,
                      theme_set(theme_bw()))

plot.sim <- function(sim.out,nd){
  data.frame(day=seq(1:(nd+1)),
             Weight=sim.out$W,
             Carapace_Width=sim.out$CW) %>%
    pivot_longer(cols=c(2:3), names_to="Metric") %>%
    ggplot(aes(x=day,y=value)) +
    geom_path() +
    facet_wrap(~Metric, scales="free") + ylab("mm / grams")
}


dge_dir <- '../../../DungiGraciisBioenergetics'
nrg_dir <- '../../../NRG'
```
```{r Part0, echo=FALSE, warning=FALSE}
  # install.packages("devtools")
  #library("devtools");devtools::install_github("kholsman/NRG")

  #library("NRG")
  #setwd("/Users/kholsman/GitHub/NRG/NRG")

setwd('C:/Users/mfisher5/Documents/NRG')
source("R/make.R")


# source("R/G_fun.R")
# source("R/bioE.R")
# source("R/load_data.R")
# source("R/cuml_curv.R")
# source("R/ftc_fun.R")
# source("R/find_ftcpar.R")
# source("R/hello.R")
# source("R/nrg.R")
# source("R/plot_curve.R")
# source("R/plot_curve3.R")
# source("R/profileT.R")
# source("R/profileW.R")
# source("R/resp_fun.R")
# source("R/rmd2md.R")
# source("R/sim_W.R")
# source("R/simul_ftc.R")
# source("R/ultils.R")
# source("R/waste_fun.R")
source("R/subscripts/molty_IP.R")
source("R/subscripts/molty_MI.R")
source("R/subscripts/crab_afdw.R")
```
<br>

# 1. Getting set up

## 1.1. Parameters

```{r}
parms <- crabPARMS_USE %>% filter(X=="Cmagister"); str(parms)
```
<br>



Note that: 

- Ceq is opt 3,

- Req is opt 0 -- user-specified respiration model. 

<br>
<br>

## 1.2. Data object 

The Dungeness crab data object is saved into NRG. Note that the units in this data / additional parameters all match the units in the parameters object. The respiration function, exuvia energy loss, and proportion I loss are from McDonald et al. (2006). Let's set prey energy density is to be equivalent to predator energy density, representing a 100% cannibalistic diet (!). 
```{r}
str(dcrab_data)

dcrab_data$Eprey <- dcrab_data$Epred
```

<br>

Molting information is also provided as a data frame.  
```{r}
str(dataMOLT$Cmagister)
```

Energy lost to exuvia is modeled as a proportion of body mass. The molt model is set to `1`, which means that molting increment will be based on carapace width (as opposed to degree-days)

<br>

### 1.3. Temperature

For now, use temperature loggers from the intertidal in Willapa Bay, WA. Provided by WSG Crab Team

```{r}
load(here(nrg_dir,'data','WillapaBayTemperature.rds'))
```


Get a daily average temperature
```{r fig.width=4, fig.height=3}
Tdat_raw <- WillapaBayTemperature$WB.oysterville.2021

Tdat <- Tdat_raw %>%
  group_by(date) %>%
  summarise(TempC=mean(TempC))
ggplot() + geom_point(data=Tdat_raw, aes(x=date, y=TempC), col="grey60") +
  geom_path(data=Tdat, aes(x=date,y=TempC, group=1), col="darkred", linewidth=2) 
```
<br>

Then subset the temperature data to run between Jul 1 and October 1.
```{r}
summerTdat <- Tdat %>% filter(ymd(date) < ymd('2021-10-01') & ymd(date) > ymd('2021-07-01'))
```


### 1.4. Test sim_W

Make sure that NRG will work with data given, by running sim_w. 

#### tiny crab

This crab is small, so it should molt frequently

Set up the simulation with some fake weights
```{r}
nd     <-  dim(summerTdat)[1] # days of simulation
# at start of simulation
Wstart <- 0.1           

## made up simulation day, weights + date 
Wobs   <- data.frame(
  day = c(1,nd),
  W   = c(Wstart,0.8)) 
Wobs$date <- summerTdat$date[Wobs$day]

# simulation data
sim_dat       <- dcrab_data
sim_dat$TempC <- summerTdat$TempC     ## temperatures
sim_dat$W     <- Wstart         ## set W data for the simulation to the first observed weight

DCRBdat4sim   <- list(
                  Wtarget  = Wobs,
                  parSIM  = parms,
                  simTdat = summerTdat,
                  dataSIM  = sim_dat,
                  dataMOLT = dataMOLT$Cmagister)

sim_out       <- summerTdat %>% mutate(W=0)
```
```{r}
dcrb.oy.sim <- sim_W(par=0.4, dataIN=DCRBdat4sim, LL=F)
```

```{r echo=FALSE}
plotdat <- data.frame(day=seq(1:(nd+1)),
                      Weight=dcrb.oy.sim$W,
                      Carapace_Width=dcrb.oy.sim$CW) %>%
  pivot_longer(cols=c(2:3), names_to="Metric")
ggplot(plotdat,aes(x=day,y=value)) +
  geom_path() +
  facet_wrap(~Metric, scales="free") + ylab("mm / grams")
```



#### mid-sized crab

This crab is a little larger, so it should molt less frequently

Set up the simulation with some fake weights
```{r}
nd     <-  dim(summerTdat)[1] # days of simulation
# at start of simulation
Wstart <- 0.8           

## made up simulation day, weights + date 
Wobs   <- data.frame(
  day = c(1,nd),
  W   = c(Wstart,1.2)) 
Wobs$date <- summerTdat$date[Wobs$day]

# simulation data
sim_dat       <- dcrab_data
sim_dat$TempC <- summerTdat$TempC     ## temperatures
sim_dat$W     <- Wstart         ## set W data for the simulation to the first observed weight

DCRBdat4sim   <- list(
                  Wtarget  = Wobs,
                  parSIM  = parms,
                  simTdat = summerTdat,
                  dataSIM  = sim_dat,
                  dataMOLT = dataMOLT$Cmagister)

sim_out       <- summerTdat %>% mutate(W=0)
```
```{r}
dcrb.oy.sim <- sim_W(par=0.4, dataIN=DCRBdat4sim, LL=F)
```

```{r echo=FALSE}
plotdat <- data.frame(day=seq(1:(nd+1)),
                      Weight=dcrb.oy.sim$W,
                      Carapace_Width=dcrb.oy.sim$CW) %>%
  pivot_longer(cols=c(2:3), names_to="Metric")
ggplot(plotdat,aes(x=day,y=value)) +
  geom_path() +
  facet_wrap(~Metric, scales="free") + ylab("mm / grams")
```


### 1.5 Crab sizes

What is the smallest sized crab that I collected? Convert this to mass (g AFDW) using the conversion equation in NRG. (*Note* - the smallest size I recorded falls within the "late cohor" J1 instar size range, which is usually present in the Puget Sound intertidal in early July)
```{r}
lab_metadat <- read_csv(here('../../','data','metadata','crab_metadat.csv'))
min(lab_metadat$CW_mm)
W_start <- cw2afdw(CW=5); W_start
```


By late September, crab in the Puget Sound are generally at the J7 stage; ~40mm carapace width is the largest size for the "late cohort" range, and the smallest size for the "early cohort" range. Convert this to mass (g AFDW) using the conversion equation in NRG. 
```{r}
W_end <- cw2afdw(CW=40); W_end
```



# 2. Prep Diet Profiles

Read in all of the prey items identified as having particularly high / low OA impacts, and their energy densities. The "High Exposure" diet will consist only of prey items with lower / positive OA impact, whereas the "Low Exposure" diet will pull from the full suite of prey taxa.
```{r}
oadiets <- read_csv(here('../../','data','bioE','OAdiets_HiLo_Condensed_EnergyDensities.csv'))

oadiets %<>%
  bind_rows(oadiets %>% filter(OA.diet.group=="high") %>% mutate(OA.diet.group="low"))
```
```{r}
oadiets %>% ggplot(aes(x=OA.diet.group,y=raw_survival_scalar)) +
  geom_boxplot()+
  geom_jitter(aes(color=functional_group), width=0.1) +
  scale_color_viridis_d(option="rainbow", name="Functional Group") +
  labs(x="OA Exposure Diet",y="Survival Scalar") +
  theme(axis.text.x=element_text(angle=45,hjust=1))
```

```{r echo=FALSE}
pdiet1 <- ggplot(oadiets, aes(x=fct_reorder(functional_group,raw_survival_scalar),y=ed_JgAFDW)) +
  geom_point(aes(color=as.factor(raw_survival_scalar))) +
  labs(x="Prey Group (Hi --> Lo Impact)",y="Energy density (J/g AFDW)") +
  scale_colour_viridis_d(option="H") +
  theme(axis.text.x=element_text(angle=45,hjust=1))
pdiet1
```

```{r}
png(here('../../','figs','figS_bioE_OAdiets_EDvGroup.png'))
pdiet1
dev.off()
```



```{r echo=FALSE}
pdiet2 <- ggplot(oadiets, aes(x=OA.diet.group,y=ed_JgAFDW)) +
  geom_boxplot()+
  geom_jitter(aes(color=functional_group), width=0.1) +
  scale_color_viridis_d(option="rainbow", name="Functional Group") +
  labs(x="OA Exposure Diet",y="Energy density (J/g AFDW)", main="All Diet Taxa") +
  theme(axis.text.x=element_text(angle=45,hjust=1))
pdiet2
```

```{r}
png(here('../../','figs','figS_bioE_OAdiets_EDvExposureGroup.png'), res=300,height=1200,width=1500)
pdiet2
dev.off()
```

Split each Exposure into high / low energy density groups. I can split pretty much down the middle if I use "greater than" the median for high OA exposure. There are enough taxa for low OA exposure, that I just grabbed the six with the highest and lowest energy densities. 
```{r}
oadiets.grouped <- oadiets %>% filter(OA.diet.group=="high") %>%
       mutate(med.ED=median(ed_JgAFDW,na.rm=TRUE)) %>%
       mutate(ED.group=ifelse(ed_JgAFDW > med.ED,"high","low")) %>% dplyr::select(-med.ED) %>%
  bind_rows(
    bind_rows(oadiets %>% filter(OA.diet.group=="low") %>%
              slice_max(order_by=ed_JgAFDW, n=5) %>% mutate(ED.group="high"),
              oadiets %>% filter(OA.diet.group=="low") %>%
              slice_min(order_by=ed_JgAFDW, n=5) %>% mutate(ED.group="low")))
oadiets.grouped %>% group_by(OA.diet.group,ED.group) %>% summarise(n())
```

**reset RFR to 1.0**
```{r}
parms$RFR <- 1
```


# 3a. Simulate Growth: 1 prey item

Start by simulating a diet composed of only 1 prey item (the highest / lowest ED prey)
```{r}
diets.df1 <- oadiets.grouped %>%
  ## highest ED prey for each group
  group_by(OA.diet.group) %>%
  slice_max(order_by=ed_JgAFDW,n=1) %>%
  ## lowest ED prey for each group
  bind_rows(oadiets.grouped %>%
              group_by(OA.diet.group) %>%
              slice_max(order_by=(-ed_JgAFDW),n=1))
```

### low OA exposure

The prey item with the highest / lowest ED, of all taxa
```{r}
tmp_diet <- diets.df1 %>% filter(OA.diet.group=="low") %>%
  dplyr::select(new_taxon2,ED.group,ed_JgAFDW)

tmp_diet
```


Set up the simulation: **high energy density**
```{r}
filter(tmp_diet, ED.group=="high")$ed_JgAFDW
```
```{r}
sim_dat <- dcrab_data

## adjust diet information
sim_dat$Eprey <- filter(tmp_diet, ED.group=="high")$ed_JgAFDW
sim_dat$propDiet <- 1

## temperatures
sim_dat$TempC <- summerTdat$TempC     ## temperatures

## starting crab weight
sim_dat$W     <- W_start         ## set W data for the simulation to the first observed weight


##
nd     <-  dim(summerTdat)[1] # days of simulation

## weights at start and end of simulation
Wobs   <- data.frame(
  day = c(1,nd),
  W   = c(W_start,W_end)) 
Wobs$date <- summerTdat$date[Wobs$day]

# simulation data frames
DCRBdat4sim   <- list(
                  Wtarget  = Wobs,
                  parSIM  = parms,
                  simTdat = summerTdat,
                  dataSIM  = sim_dat,
                  dataMOLT = dataMOLT$Cmagister)

sim_out       <- summerTdat %>% mutate(W=0)
```

simulate!
```{r}
oa1.lowOhiE.sim <- sim_W(par=parms$RFR, dataIN=DCRBdat4sim, LL=F)
plot.sim(sim.out =oa1.lowOhiE.sim,nd=nd)
```

The carapace widths look normal, but that weight is out of control!
final weight:
```{r}
oa1.lowOhiE.sim$W[nd]
```



Set up the simulation: **low energy density**
```{r}
filter(tmp_diet, ED.group=="low")$ed_JgAFDW
```
```{r}
sim_dat <- dcrab_data

## adjust diet information
sim_dat$Eprey <- filter(tmp_diet, ED.group=="low")$ed_JgAFDW
sim_dat$propDiet <- 1

## temperatures
sim_dat$TempC <- summerTdat$TempC     ## temperatures

## starting crab weight
sim_dat$W     <- W_start         ## set W data for the simulation to the first observed weight


##
nd     <-  dim(summerTdat)[1] # days of simulation

## weights at start and end of simulation
Wobs   <- data.frame(
  day = c(1,nd),
  W   = c(W_start,W_end)) 
Wobs$date <- summerTdat$date[Wobs$day]

# simulation data frames
DCRBdat4sim   <- list(
                  Wtarget  = Wobs,
                  parSIM  = parms,
                  simTdat = summerTdat,
                  dataSIM  = sim_dat,
                  dataMOLT = dataMOLT$Cmagister)

sim_out       <- summerTdat %>% mutate(W=0)
```

simulate!
```{r}
oa1.lowOlowE.sim <- sim_W(par=parms$RFR, dataIN=DCRBdat4sim, LL=F)
plot.sim(sim.out =oa1.lowOlowE.sim,nd=nd)
```

final weight:
```{r}
oa1.lowOlowE.sim$W[nd]
```


How does the "high OA exposure" diet this compare to the  "low OA exposure" diet in terms of...

Growth each day
```{r fig.height=3,fig.width=6}
p1 <- data.frame(TempC=summerTdat$TempC,
                        date=summerTdat$date,
                        day_of_sim=seq(1,nd),
                        ED_hi=oa1.lowOhiE.sim$G_ggd[1:nd],
                        ED_low=oa1.lowOlowE.sim$G_ggd[1:nd]) %>%
  pivot_longer(cols=c(4:5),names_to="Simulation", values_to="G_ggd") %>%
ggplot( aes(x=day_of_sim,y=G_ggd,col=Simulation)) +
  geom_point(size=0.5) +
  geom_path() +
  geom_hline(aes(yintercept=0)) +
  labs(title="low OA exposure") + theme(legend.position = "none")
p2 <- data.frame(TempC=summerTdat$TempC,
                        date=summerTdat$date,
                        day_of_sim=seq(1,nd),
                        ED_hi=oa1.lowOhiE.sim$W[1:nd],
                        ED_low=oa1.lowOlowE.sim$W[1:nd]) %>%
  pivot_longer(cols=c(4:5),names_to="Simulation", values_to="Weight (g AFDW)") %>%
ggplot( aes(x=date,y=`Weight (g AFDW)`,col=Simulation)) +
  geom_point(size=0.5) +
  geom_path() +
  geom_hline(aes(yintercept=0)) +
  labs(title="")
plot_grid(p1,p2,ncol=2, rel_widths=c(0.7,1))
```



### high OA exposure

The prey item with the highest / lowest ED, of all taxa
```{r}
tmp_diet <- diets.df1 %>% filter(OA.diet.group=="high") %>%
  dplyr::select(new_taxon2,ED.group,ed_JgAFDW)

tmp_diet
```


Set up the simulation: **high energy density**
```{r}
filter(tmp_diet, ED.group=="high")$ed_JgAFDW
```
```{r}
sim_dat <- dcrab_data

## adjust diet information
sim_dat$Eprey <- filter(tmp_diet, ED.group=="high")$ed_JgAFDW
sim_dat$propDiet <- 1

## temperatures
sim_dat$TempC <- summerTdat$TempC     ## temperatures

## starting crab weight
sim_dat$W     <- W_start         ## set W data for the simulation to the first observed weight


##
nd     <-  dim(summerTdat)[1] # days of simulation

## weights at start and end of simulation
Wobs   <- data.frame(
  day = c(1,nd),
  W   = c(W_start,W_end)) 
Wobs$date <- summerTdat$date[Wobs$day]

# simulation data frames
DCRBdat4sim   <- list(
                  Wtarget  = Wobs,
                  parSIM  = parms,
                  simTdat = summerTdat,
                  dataSIM  = sim_dat,
                  dataMOLT = dataMOLT$Cmagister)

sim_out       <- summerTdat %>% mutate(W=0)
```

simulate!
```{r}
oa1.hiOhiE.sim <- sim_W(par=parms$RFR, dataIN=DCRBdat4sim, LL=F)
plot.sim(sim.out =oa1.hiOhiE.sim,nd=nd)
```

final weight:
```{r}
oa1.hiOhiE.sim$W[nd]
```



Set up the simulation: **low energy density**
```{r}
filter(tmp_diet, ED.group=="low")$ed_JgAFDW
```
```{r}
sim_dat <- dcrab_data

## adjust diet information
sim_dat$Eprey <- filter(tmp_diet, ED.group=="low")$ed_JgAFDW
sim_dat$propDiet <- 1

## temperatures
sim_dat$TempC <- summerTdat$TempC     ## temperatures

## starting crab weight
sim_dat$W     <- W_start         ## set W data for the simulation to the first observed weight


##
nd     <-  dim(summerTdat)[1] # days of simulation

## weights at start and end of simulation
Wobs   <- data.frame(
  day = c(1,nd),
  W   = c(W_start,W_end)) 
Wobs$date <- summerTdat$date[Wobs$day]

# simulation data frames
DCRBdat4sim   <- list(
                  Wtarget  = Wobs,
                  parSIM  = parms,
                  simTdat = summerTdat,
                  dataSIM  = sim_dat,
                  dataMOLT = dataMOLT$Cmagister)

sim_out       <- summerTdat %>% mutate(W=0)
```

simulate!
```{r}
oa1.hiOlowE.sim <- sim_W(par=parms$RFR, dataIN=DCRBdat4sim, LL=F)
plot.sim(sim.out =oa1.hiOlowE.sim,nd=nd)
```

final weight:
```{r}
oa1.hiOlowE.sim$W[nd]
```







### compare across OA exposures

With the highest energy density possible, a crab with high OA exposure will have what weight compared to a crab with low OA exposure? *a 33% decline in weight from low to high exposure*
```{r}
w.lo <- oa1.lowOhiE.sim$W[nd]
w.hi <- oa1.hiOhiE.sim$W[nd]

w.hi/w.lo
(w.hi - w.lo)/w.lo * 100
```


With the lowest energy density possible, a crab with high OA exposure will have what weight compared to a crab with low OA exposure? *equal (no change)*
```{r}
w.lo <- oa1.lowOlowE.sim$W[nd]
w.hi <- oa1.hiOlowE.sim$W[nd]


w.hi/w.lo
(w.hi - w.lo)/w.lo * 100
```



# 3b. Simulate Growth: multiple prey items


randomly sample a set of 6 numbers from a normal distribution, and normalize to add to "1". do this 1000 x. These will be the diet proportions. 
```{r}

```
















