---
title: "Creating Model Diets"
author: "M Fisher"
date: "2023-09-05"
output: html_document
---

# Description

From the dDNA data, summarize model diets for two analyses: (1) potential site-specific growth differences associated with different diets, (2) potential growth differences associated with diet item OA sensitivity.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(vegan)
library(magrittr)
library(janitor)

# User inputs
blastdir   <- 'data/blast'
bdir <- 'data/bioE'
run.nums <- c(1,2)
marker  <- 'lerayXT'
```

```{r data, include=FALSE}
## filtered taxa ##
dat <- read_csv(here(blastdir, paste0(marker,"_","r",paste(run.nums,collapse="-"),"_sample_taxonomy_filtered2_uniqueTaxa.csv")))

dat %<>%
  mutate(site=ifelse(site=="KAY" & crab_num > 18, "KAY-shell",site)) %>%
  mutate(site=ifelse(site=="CLAY","LAR",
                     ifelse(site=="SIN","SAMI",site))) %>%
  mutate(estuary=ifelse(site %in% c("KAY","KAY-shell"),"Port Susan Bay",
                                    ifelse(site %in% c("LAR","SAMT","SAMI"), "Samish Bay","Padilla Bay")))

## too small taxa (aka "detritus") ##
smdat <- read_csv(here(blastdir, paste0(marker,"_","r",paste(run.nums,collapse="-"),"_sample_taxonomy_SmallTaxa_uniqueTaxa.csv")))  

smdat %<>%
  mutate(site=ifelse(site=="KAY" & crab_num > 18, "KAY-shell",site)) %>%
  mutate(site=ifelse(site=="CLAY","LAR",
                     ifelse(site=="SIN","SAMI",site))) %>%
  mutate(estuary=ifelse(site %in% c("KAY","KAY-shell"),"Port Susan Bay",
                                    ifelse(site %in% c("LAR","SAMT","SAMI"), "Samish Bay","Padilla Bay")))

alldat <- bind_rows(dat,smdat)
```
<br>

## Basics

What is the median taxonomic richness per crab?
with detritus
```{r}
median(bind_rows(dat,smdat) %>%
  group_by(crab_id) %>%
  summarise(richness=length(unique(taxon))) %>% pull(richness))
```
with detritus as one category ("uni-cellular")
```{r}
median(bind_rows(dat,smdat %>% mutate(taxon="unicellular")) %>%
  group_by(crab_id) %>%
  summarise(richness=length(unique(taxon))) %>% pull(richness))
```

without detritus
```{r}
median(dat %>%
  group_by(crab_id) %>%
  summarise(richness=length(unique(taxon))) %>% pull(richness))
```
only detritus
```{r}
median(smdat %>%
  group_by(crab_id) %>%
  summarise(richness=length(unique(taxon))) %>% pull(richness))
```

What is the median stomach fullness per crab?
```{r}
```


## Overall diet





## OA sensitivity diets

Produce four diets: Hi/Lo OA Sensitivity x Hi/Lo Energy Density

First, grab only the "large" prey that were present in three or more crab
```{r}
summary_dat <- dat %>% 
  group_by(group, new_taxon) %>% summarise(ncrab=length(unique(crab_id)))

dat_3plus <- dat %>%
  filter(new_taxon %in% filter(summary_dat, ncrab > 2)$new_taxon)

## how many prey?
length(unique(dat_3plus$new_taxon))

## how many crab? 
length(unique(dat_3plus$crab_id))

write_csv(summary_dat %>% filter(ncrab > 2), here(bdir,'OAdiets_all_taxa_3plus_crab.csv'))

```

Grab "small" prey that were present in three or more crab.
```{r}
summary_smdat <- smdat %>% 
  group_by(phylum,taxon) %>% summarise(ncrab=length(unique(crab_id)))

smdat_3plus <- smdat %>%
  filter(taxon %in% filter(summary_smdat, ncrab > 2)$taxon)

## how many prey?
length(unique(smdat_3plus$taxon))

## how many crab? 
length(unique(smdat_3plus$crab_id))

write_csv(summary_smdat %>% filter(ncrab > 2), here(bdir,'OAdiets_small_taxa_3plus_crab.csv'))

```


Then, match these prey to the OA list, according to the manually identified functional groups. 

Uncertainties in manual matching: 

- Microzooplankton the agreement score is low. This seems to be because some do make caco3, and some don’t. Could potentially fine-tune this value according to species-level IDs in stomach content data?

- Tunicates (Botrylloides) and rotifers (Euchlanis dilatata) not in the species database. Euchlanis dilatata is a tofier that is generally 170-250uM in length. Compare to Alexandrium, a genus of dinoflagellites classified as microzooplankton at about 50uM, and the copepod Paracalanus, classified as a mesozooplankton at 0.5-1mm (500uM). Euclanis dilatata therefore classified as a microzooplankton. Green Colonial Tunicate was listed as a Shallow benthic filter feeder, therefore Botrylloides as that functional group.


```{r}
taxdat <- read_csv(here(bdir,'OAdiets_all_taxa_3plus_crab_OAinfo.csv'))
oadat <- read_csv(here(bdir,'BuschMcElhany2016_OAscores_by_functionalGroup.csv'),skip=1) %>% clean_names()

colnames(oadat)
```

```{r}
taxdat %<>% left_join(oadat, by="functional_group")
```

Check the raw survival scalar frequencies
```{r}
ggplot(taxdat, aes(x=raw_survival_scalar)) +
  geom_histogram() + 
  geom_vline(aes(xintercept=mean(taxdat$raw_survival_scalar))) +
  geom_text(aes(x=mean(taxdat$raw_survival_scalar), y=5, label=round(mean(taxdat$raw_survival_scalar),2))) +
  theme_bw()
```


What is 1 standard deviation away from the mean? how many taxa are above / below that cutoff?
```{r}
xbar <- mean(taxdat$raw_survival_scalar)
sdev <- sd(taxdat$raw_survival_scalar)
upper <- xbar + sdev; lower <- xbar - sdev

upper
taxdat %>% filter(raw_survival_scalar > upper) %>% pull(new_taxon)

lower
taxdat %>% filter(raw_survival_scalar < lower) %>% pull(new_taxon)
```

what is the 50% IQR? how many taxa are above / below that cutoff?
```{r}
xmed <- median(taxdat$raw_survival_scalar)
q75 <- quantile(taxdat$raw_survival_scalar, probs=0.75)
q25 <- quantile(taxdat$raw_survival_scalar, probs=0.25)
xmed

q75
taxdat %>% filter(raw_survival_scalar > q75) %>% pull(new_taxon)

q25
taxdat %>% filter(raw_survival_scalar < q25) %>% pull(new_taxon)
```
Write these out!
```{r}
write_csv(taxdat, here(bdir,'OAdiets_all_taxa_3plus_crab_OAinfo.csv'))

diets_out <- taxdat %>% mutate(OA.diet.group=ifelse(raw_survival_scalar < q25, "low",
                                                    ifelse(raw_survival_scalar > q75, "high",NA))) %>%
  filter(!is.na(OA.diet.group)) %>%
  dplyr::select(new_taxon,OA.diet.group,colnames(taxdat)[-1])
write_csv(diets_out, here(bdir,'OAdiets_HiLo.csv'))
```






