---
title: "Calculate OA Risk from diet"
author: "M Fisher"
date: "2023-09-05"
output: html_document
---

# Description

From the dDNA data, attach (a) OA impact information, and (b) energy densities, to each prey item. Use this in combination with frequency of occurrence to evaluate risk.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(vegan)
library(magrittr)
library(janitor)

# User inputs
blastdir   <- '../../data/blast'
resultsdir <- '../../data/results'
bdir       <- '../../data/bioE'
run.nums <- c(1,2)
marker  <- 'lerayXT'
knitting=FALSE
```

```{r data, include=FALSE}
## filtered taxa ##
dat <- read_csv(here(blastdir, paste0(marker,"_","r",paste(run.nums,collapse="-"),"_sample_taxonomy_filtered2_uniqueTaxa.csv")))

dat %<>%
  mutate(site=ifelse(site=="KAY" & crab_num > 18, "KAY-shell",site)) %>%
  mutate(site=ifelse(site=="CLAY","LAR",
                     ifelse(site=="SIN","SAMI",site))) %>%
  mutate(estuary=ifelse(site %in% c("KAY","KAY-shell"),"Port Susan Bay",
                                    ifelse(site %in% c("LAR","SAMT","SAMI"), "Samish Bay","Padilla Bay")))

## too small taxa (aka "detritus") ##
smdat <- read_csv(here(blastdir, paste0(marker,"_","r",paste(run.nums,collapse="-"),"_sample_taxonomy_SmallTaxa_uniqueTaxa.csv")))  

smdat %<>%
  mutate(site=ifelse(site=="KAY" & crab_num > 18, "KAY-shell",site)) %>%
  mutate(site=ifelse(site=="CLAY","LAR",
                     ifelse(site=="SIN","SAMI",site))) %>%
  mutate(estuary=ifelse(site %in% c("KAY","KAY-shell"),"Port Susan Bay",
                                    ifelse(site %in% c("LAR","SAMT","SAMI"), "Samish Bay","Padilla Bay")))

## carapace width
cw.df <- readRDS(here('../../','data','metadata','carapace_widths_for_zoid.rds'))
```
```{r eval=FALSE}
## may have to save all 
summary_rare_dat <- bind_rows(dat, mutate(smdat, new_taxon=taxon)) %>% 
  mutate(group=ifelse(is.na(group) & is.na(phylum),class,
                      ifelse(is.na(group) & !is.na(phylum),phylum,group))) %>%
  group_by(group, new_taxon) %>% summarise(ncrab=length(unique(crab_id))) %>%
  filter(ncrab < 3) %>% arrange(ncrab)

write_csv(summary_rare_dat, here(bdir, 'OAdiets_all_rare_taxa_1-2crab.csv'))
```
<br>

# Crab Exposure Index

## Prey item frequency

Calculate the frequency of occurrence for all prey items, across all crabs. *I've already done this in a previous script.*
```{r}
prey.fo <- read_csv( here(resultsdir,'freqoccur_runs1-2_lerayxt_presence.csv'))
```

Calculate the frequency of occurrence for prey items, across crabs within a given estuary.
```{r}
prey.fo.estuary <- bind_rows(dat, mutate(smdat, new_taxon=taxon)) %>% 
  group_by(estuary) %>% mutate(ncrab.estuary=length(unique(crab_id))) %>%
  ungroup() %>%
  group_by(estuary, new_taxon, ncrab.estuary) %>%
  summarise(ncrab=length(unique(crab_id))) %>%
  mutate(pcrab=(ncrab/ncrab.estuary))

```


## OA impact to prey

Match all prey to the OA list, according to the manually identified functional groups. 

```{r}
taxdat <- read_csv( here(bdir, 'OAdiets_all_taxa_toMatch.csv'))

# how many functional groups are missing?
with(filter(taxdat,is.na(functional_group)), table(ncrab))
```



Uncertainties in manual matching: 

- Microzooplankton the agreement score is low. This seems to be because some do make caco3, and some don’t. Could potentially fine-tune this value according to species-level IDs in stomach content data?

- Tunicates (Botrylloides) and rotifers (Euchlanis dilatata) not in the species database. Euchlanis dilatata is a tofier that is generally 170-250uM in length. Compare to Alexandrium, a genus of dinoflagellites classified as microzooplankton at about 50uM, and the copepod Paracalanus, classified as a mesozooplankton at 0.5-1mm (500uM). Euclanis dilatata therefore classified as a microzooplankton. Green Colonial Tunicate was listed as a Shallow benthic filter feeder, therefore Botrylloides as that functional group.

- Nematodes not in the species database ("Nematode community" in species key, but no family reference to link to functional group). But listed as meiobenthos in functional groups.

- Grouped free-living platyhelminthes in with nematodes, in the meiobenthos

- Snails that were facultative carnivores / omnivores did not strictly fit into the category "Benthic herbivorous grazers" or "Sea stars moonsnail whelk", so they were conservatively put into the category with the lowest OA impact.  (n=1 taxa in n=1 crab)


```{r}
oadat <- read_csv(here(bdir,'BuschMcElhany2016_OAscores_by_functionalGroup.csv'),skip=1) %>% clean_names()
colnames(oadat)
```

```{r}
if(!knitting){
  
  taxdat %<>% filter(!is.na(functional_group)) %>% left_join(oadat, by="functional_group")
  sum(is.na(taxdat$raw_survival_scalar))
  
}
```

Write these out!
```{r echo=TRUE}
if(!knitting){
  
write_csv(taxdat, here(bdir,'OAimpact_runs1-2_lerayxt_allTaxa.csv'))

}
```


## Crab Index of Exposure: Prey Exposure x Freq

### all data

Calculate the crab index of exposure for each prey item $i$ from the Euclidean distance between $x_i$ = prey item frequency of occurrance (`pcrab`) and $y_i$ = OA impact on prey item (`relative survival scalar`)
```{r}
eindex.dat <- taxdat %>% left_join(prey.fo,by=c("new_taxon"="taxon")) %>%
  mutate(pcrab=pcrab/100) %>%
  mutate(std_survival_scalar=decostand(x=matrix(data=taxdat$raw_survival_scalar, ncol=1), method="range", MARGIN=2)) %>%
  mutate(eindex=sqrt(pcrab^2 + std_survival_scalar^2))

eindex.dat %<>% mutate(eindex.rescaled=decostand(x=matrix(data=eindex.dat$eindex, ncol=1), method="range", MARGIN=2))
```

```{r}
p1.eindex <- eindex.dat %>%
  ggplot(aes(x=pcrab,y=std_survival_scalar, col=eindex.rescaled)) +
  geom_point(size=2) +
  labs(x="Frequency of Prey Occurrence", y="OA Impact on Prey") +
  xlim(c(0,1)) +
  scale_color_distiller(palette="PuRd", name="Crab Index\nof Exposure", direction=1) 
  # scale_color_viridis_c(option="H",begin=0.1, end=0.8, name="Crab Index\nof Exposure")
p1.eindex
```

Remove points that lie above the 99th percentile
```{r}
fo.max <- as.numeric(quantile(x=eindex.dat$pcrab, probs=0.99))
p2.eindex <- eindex.dat %>%
  filter(pcrab < fo.max) %>%
  ggplot(aes(x=pcrab,y=std_survival_scalar, col=eindex.rescaled)) +
  geom_point(size=2) +
  labs(x="Frequency of Prey Occurrence", y="OA Impact on Prey") +
  xlim(c(0,0.4)) +
  scale_color_distiller(palette="PuRd", name="Crab Index\nof Exposure", direction=1) 
  # scale_color_viridis_c(option="H",begin=0.1, end=0.8, name="Crab Index\nof Exposure")
p2.eindex
```

for each estuary
```{r}
eindexByEstuary <- taxdat %>% left_join(prey.fo.estuary,by=c("new_taxon"))
eindexByEstuary %<>%
  mutate(std_survival_scalar=decostand(x=matrix(data=eindexByEstuary$raw_survival_scalar, ncol=1), method="range", MARGIN=2)) %>%
  mutate(eindex=sqrt(pcrab^2 + std_survival_scalar^2))
eindexByEstuary %<>% mutate(eindex.rescaled=decostand(x=matrix(data=eindexByEstuary$eindex, ncol=1), method="range", MARGIN=2))
```
```{r}
p.eindex3 <- eindexByEstuary %>%
  ggplot(aes(x=pcrab,y=std_survival_scalar, col=eindex.rescaled)) +
  geom_point(size=2) +
  facet_grid(cols=vars(estuary)) +
  labs(x="Frequency of Prey Occurrence", y="OA Impact on Prey") +
  xlim(c(0,1)) +
  scale_color_distiller(palette="PuRd", name="Crab Index\nof Exposure", direction=1) 
  # scale_color_viridis_c(option="H",begin=0.1, end=0.8, name="Crab Index\nof Exposure")
p.eindex3 
```

save the plot. first for all crab:
```{r}
png(here('../../','figs','fig5a_crab_eindex_all.png'),res=300,height=2000,width=2300)

ggdraw(p2.eindex) +
  draw_plot(p1.eindex + 
              scale_x_continuous(limits=c(0,1),breaks=c(0,0.5,1),labels=c(0,0.5,1)) +
              scale_y_continuous(limits=c(0,1),breaks=c(0,0.5,1),labels=c(0,0.5,1)) +
              theme(axis.title=element_blank(),legend.position="none"), 
            x=.54, y=0.08, width=0.3, height=.3)

dev.off()
```

and then by estuary:
```{r}
png(here('../../','figs','figS_crab_OAexposure_index_byEstuary.png'),res=300,height=1200,width=2200)
p.eindex3 + theme(axis.text.x=element_text(angle=25, hjust=1))
dev.off()
```


## Crab Risk: Exposure Index x Sensitivity

Combine the exposure index above ($x$) with crab sensitivity to get risk. Sensitivity is defined as the energy density of the prey item. 

I don't currently have all energy densities, just those for the prey items with the highest / lowest OA impact scalars. 
```{r}
ed.dat <- read_csv(here(bdir, 'OAdiets_HiLo_energy_Densities.csv')) %>% dplyr::select(-OA.diet.group)
```

```{r}
risk.dat <- eindex.dat %>%
  left_join(ed.dat %>% dplyr::select(-functional_group,-raw_survival_scalar,-ncrab), by=c("new_taxon")) %>%
  filter(!is.na(ed_JgAFDW))

# standardize energy densities
risk.dat %<>% mutate(std_ed=decostand(matrix(risk.dat$ed_JgAFDW,ncol=1),method="range",MARGIN=2))
```

Calculate overall risk
```{r}
risk.dat %<>% mutate(crab.risk=sqrt(eindex.rescaled^2 + std_ed^2))

risk.dat %<>% mutate(crab.risk=as.numeric(decostand(x=matrix(data=risk.dat$crab.risk, ncol=1), method="range", MARGIN=2)))
```


```{r}
# p1.eindex <- 
risk.dat %>%
  ggplot(aes(x=std_survival_scalar,y=std_ed, color=crab.risk)) +
  geom_point(size=2) +
  geom_text_repel(aes(label=new_taxon), fontface='italic', min.segment.length = 0.01) +
  scale_color_viridis_c(option="H",begin=0.4, end=0.8, name="Crab Risk") +
  labs(x="Crab Exposure Index", y="Crab Sensitivity")
# p1.eindex
```

```{r}
risk.dat %<>% mutate(std_survival_scalar=as.numeric(std_survival_scalar),
                     eindex=as.numeric(eindex),
                     eindex.rescaled=as.numeric(eindex.rescaled),
                     std_ed=as.numeric(std_ed))
write_csv(risk.dat, here(bdir,'OArisk_runs1-2_lerayxt_2023-10-06.csv'))
```

```{r}

png(here('../../','figs','fig5b_crab_OArisk_all.png'),res=300,height=1800,width=2100)

risk.dat %>%
  ggplot(aes(x=std_survival_scalar,y=std_ed, color=crab.risk)) +
  geom_point(size=2) +
  geom_text_repel(aes(label=new_taxon), fontface='italic', min.segment.length = 0.01) +
  scale_color_viridis_c(option="H",begin=0.4, end=0.8, name="Crab Risk") +
  labs(x="Crab Exposure Index", y="Crab Sensitivity")

dev.off()
```
