---
title: "Sampling Data"
author: "Mary Fisher"
output: 
  html_document:
      toc: yes
      toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(here)
library(tidyverse)
library(ggplot2)
library(ggmap)
library(ggspatial)
library(ggrepel)
library(janitor)
library(vegan)
library(labdsv)
library(scales)
library(ggfortify)
```

# Map of Sampling Sites

```{r include=FALSE}
# read in data
dat <- read.csv(here('data','metadata','sampling_data.csv'))

# basemap
base.map <- readRDS(here('data','metadata',"PugetSoundN_terrain_background.rds"))
wa.map <- map_data("state") %>%
  filter(region=="washington")


# colors for map points
cl <- c("darkgreen","darkturquoise","darkgoldenrod4")
```
<br>
```{r}
dat[,c(1,2,6)]

dat <- dat %>% unite("Site_N", c(Site,N), remove=FALSE)
```
<br>

```{r fig.width=10, fig.height=10}
ggmap(base.map) +
  geom_point(data=dat, aes(x=Longitude,y=Latitude, col=Habitat), stroke=3, size=4, pch=1) + 
  geom_label_repel(data=dat, aes(x=Longitude,y=Latitude, label=Site,segment.colour=Habitat),
                   force=1, force_pull=0.25, segment.size=1, size=4, min.segment.length=0.1, show.legend=FALSE) +
  ylab("Latitude") + xlab("Longitude") +
  scale_color_manual(aesthetics=c("color","segment.color"), values=cl) +
  theme(axis.title=element_text(size=14),
        axis.text=element_text(size=12),
        legend.title=element_text(size=14),
        legend.text=element_text(size=12)) +
  coord_cartesian(xlim = c(-122.9, -122), ylim = c(48, 48.75))
```
<br>

```{r fig.width=10, fig.height=10}
dat.part <- filter(dat, Site %in% c("Kayak Point","March Point", "Clayton Beach"))
ggmap(base.map) +
  geom_point(data=dat.part, aes(x=Longitude,y=Latitude, col=Habitat), stroke=3, size=4, pch=1) + 
  geom_label_repel(data=dat.part, aes(x=Longitude,y=Latitude, label=Site,segment.colour=Habitat),
                   force=1, force_pull=0.25, segment.size=1, size=4, min.segment.length=0.1, show.legend=FALSE) +
  ylab("Latitude") + xlab("Longitude") +
  scale_color_manual(aesthetics=c("color","segment.color"), values=cl) +
  theme(axis.title=element_text(size=14),
        axis.text=element_text(size=12),
        legend.title=element_text(size=14),
        legend.text=element_text(size=12)) +
  coord_cartesian(xlim = c(-122.9, -122), ylim = c(48, 48.75))
```
<br>
*The number after the sites is the sample size!*
<br>
<br>

# Group Sites by Habitat

Based on 30m transects (5 quadrats per transect at 0m, 7.5m, 15m, 22.5m, 30m)

Habitat data looks at: 

1. substrate % cover -- of sand, cobble, and shell / shell hash. total % cover per quadrat adds to 100%
2. *Z marina* and algae % cover -- the number of 5x5 squares in the quadrat that contain one or more rooted eelgrass shoots / algae. converted to a % by dividing by 25 (total number of squares)
3. *Z marina* shoot density -- the number of rooted shoots per square, for the 5 diagonal squares in the quadrat. lower bound of zero, no upper bound. 

I've only finished this for a handful of the sites in the map above. 
```{r include=FALSE}
dat <- read.csv(here('data','eelgrass_survey_data.csv')) %>% clean_names()
```
```{r}
dat[,1:3]
```
<br>


## NMDS

#### Substrate + Vegetation

By quadrat
```{r}
dat_mat <- as.matrix(dat[,4:33])
```
```{r echo=TRUE}
all.nmds <- metaMDS(dat_mat,k=2,trymax=100)
```
<br>

By average / coefficient of variation
```{r}
dat_mat <- as.matrix(dat[,34:42])
```
```{r echo=TRUE}
all.nmds <- metaMDS(dat_mat,k=2,trymax=100)
```
<br>


## PCA

The % cover values for substrate, eelgrass, and algae are all percentages (ranging from 0 to 1). However, the density data needs to be rescaled.
```{r echo=TRUE}
density_rescaled <- dat[,grep("density",colnames(dat))] %>%
  dplyr::select(-cv_marina_density) %>%
  mutate(across(.fns=rescale, to = c(0, 1)))
```
```{r}
dat_rescaled     <- dat %>%
  dplyr::select(-c(grep("density",colnames(dat)))) %>%
  bind_cols(density_rescaled) %>%
  bind_cols(dplyr::select(dat,cv_marina_density)) %>%
  dplyr::select(colnames(dat))
```
<br>

#### Substrate + Vegetation

By quadrat
```{r}
dat_mat <- as.matrix(dat_rescaled[,4:33])
```
```{r echo=TRUE}
alldat.pca <- prcomp(x=dat_mat)
```
```{r}
autoplot(alldat.pca, data = dat, colour = 'site', size=5,
         loadings = TRUE, loadings.colour = 'blue',
         loadings.label = TRUE, loadings.label.size = 2, main="Substrate/Vegetation Data") + theme_bw()
```
<br>

By average / coefficient of variation
```{r}
dat_mat <- as.matrix(dat_rescaled[,c(34:42)])
```
```{r echo=TRUE}
alldat.means.pca <- prcomp(x=dat_mat)
```
```{r}
autoplot(alldat.means.pca, data = dat, colour = 'site', size=5,
         loadings = TRUE, loadings.colour = 'blue',
         loadings.label = TRUE, loadings.label.size = 2, main="Substrate/Vegetation Summary Stats") + theme_bw()
```
<br>


#### Substrate only

```{r}
sdat_mat <- as.matrix(dat[,4:18])
substrate.pca <- prcomp(x=sdat_mat)

autoplot(substrate.pca, data = dat, colour = 'site', size=5,
         loadings = TRUE, loadings.colour = 'blue',
         loadings.label = TRUE, loadings.label.size = 2, main="Substrate, all data") + theme_bw()
```
<br>
**Not Included: the Padilla Bay NERR Sites are most similar to Clayton Beach with respect to substrate type (mostly sand, sandy mud).**
<br>

```{r}
sdat_mat <- as.matrix(dat[,34:36])
substrate.pca <- prcomp(x=sdat_mat)

autoplot(substrate.pca, data = dat, colour = 'site', size=5,
         loadings = TRUE, loadings.colour = 'blue',
         loadings.label = TRUE, loadings.label.size = 2, main="Substrate, summary stats") + theme_bw()
```

<br>


#### Vegetation only
```{r}
edat_mat <- as.matrix(dat_rescaled[,19:28])
eelgrass.pca <- prcomp(x=edat_mat)

autoplot(eelgrass.pca, data = dat, colour = 'site', size=5,
         loadings = TRUE, loadings.colour = 'blue',
         loadings.label = TRUE, loadings.label.size = 2, main="Eelgrass / Algae, all data") + theme_bw()
```
<br>

Not Included: The Padilla Bay NERR Sites are most similar to Clayton Beach with respect to eelgrass density (high) and algae cover (low to zero).
<br>

```{r}
edat_mat <- as.matrix(dat_rescaled[,37:42])
eelgrass.pca <- prcomp(x=edat_mat)

autoplot(eelgrass.pca, data = dat, colour = 'site', size=5,
         loadings = TRUE, loadings.colour = 'blue',
         loadings.label = TRUE, loadings.label.size = 2, main="Eelgrass / Algae, summary stats") + theme_bw()
```

