---
title: "PCRG Feb Updates"
author: "M Fisher"
date: "2023-02-01"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(vegan)
library(magrittr)

# User inputs
blastdir   <- 'data/blast'
resultsdir <- 'data/results'
run.nums <- c(1)
marker  <- 'lerayXT'
```
```{r data, include=FALSE}
## data from blast ##
for(i in seq(1,length(run.nums))) {
if(i==1){
    blastdat <- read_csv(here(blastdir, paste0(marker,"_","r",run.nums[i],"_sample_taxonomy_filtered.csv")))
  } else{
    blastdat <- bind_rows(blastdat,
                          read_csv(here(blastdir, paste0(marker,"_","r",run.nums[i],"_sample_taxonomy_filtered.csv"))))
  }
}

## filtered species with common names ##
for(i in seq(1,length(run.nums))) {
if(i==1){
    spdict <- read_csv(here(resultsdir, paste0(marker,"_","r",run.nums[i],"_species_filtered_common_names.csv")))
  } else{
    spdict <- bind_rows(spdict,
                          read_csv(here(resultsdir, paste0(marker,"_","r",run.nums[i],"_species_filtered_common_names.csv"))))
  }
}

dat <- blastdat %>%
  filter(taxon %in% spdict$taxon) %>%
  mutate(Site=ifelse(Site=="KAY" & Sample_num > 18, "KAY-shell",Site)) %>%
  mutate(Site=ifelse(Site=="CLAY","LAR",Site)) %>%
  left_join(spdict)
```
<br>

What are the final sample sizes for this data set?
```{r}
dat %>%
  group_by(Site) %>% summarise(`Crab count`=length(unique(crab_id)))
```


# Prey Diversity

### Alpha div

What is the species richness ($\alpha$ diversity) at each site? Richness per crab is simply the species richness / total crab from site.
```{r}
richness.site <- dat %>%
  group_by(Site) %>%
  summarise(ncrab=length(unique(crab_id)),
            species.richness=length(unique(taxon)))
richness.site
```

How is ($\alpha$ diversity) distributed across the taxa?
```{r message=FALSE}
richness.phyla.site <- dat %>%
  filter(!is.na(phylum)) %>%
  group_by(Site, phylum) %>%
  summarise(species.richness=length(unique(taxon)))  %>%
  mutate(Site=ifelse(Site=="LAR","Larrabee",ifelse(Site=="KAY","Kayak Point 1",
                                                    ifelse(Site=="KAY-shell","Kayak Point 2",
                                                           ifelse(Site=="MARPT","March Point")))))
richness.phyla <- dat %>%
  filter(!is.na(phylum)) %>%
  group_by(phylum) %>%
  summarise(species.richness=length(unique(taxon))) %>%
  arrange(species.richness)

porder <- unique(richness.phyla$phylum)

richness.phyla <- richness.phyla %>%
  mutate(Site="Overall") %>%
  bind_rows(richness.phyla.site)

richness.phyla$phylum <- factor(richness.phyla$phylum, levels=porder)
```
```{r fig.height=7,fig.width=5}
tmplot <- ggplot(richness.phyla, aes(y=phylum,x=species.richness, fill=Site)) +
  geom_col(position="dodge") +
  labs(x="Species Richness",y="Phylum") +
  scale_fill_manual(values=c("cyan4","darkblue","darkorange","coral3","gray30")) +
  theme_bw() + theme(panel.grid.major.y=element_blank())
tmplot

png(here('figs','pcrg_phyla-all-crab.png'), res=200, height=800,width=800)
tmplot
dev.off()
```


### Alpha div x Stomach Fullness
```{r message=FALSE}
richness.crab <- dat %>%
  group_by(Site,crab_id) %>%
  summarise(species.richness=length(unique(taxon))) %>% 
  mutate(Site=ifelse(Site=="LAR","Larrabee",ifelse(Site=="KAY","Kayak \nPoint 1",
                                                    ifelse(Site=="KAY-shell","Kayak \nPoint 2",
                                                           ifelse(Site=="MARPT","March \nPoint")))))
```
```{r message=FALSE}
### stomach fullness data ###
crabdat <- read_csv(here('data','metadata','crab_metadat.csv')) %>%
  filter(sample_id %in% dat$crab_id) %>%
  rename(crab_id=sample_id)

richness.crab <- richness.crab %>%
  left_join(crabdat) %>%
  filter(!is.na(stomach_fullness)) %>% filter(stomach_fullness != "0")
  
```
```{r}
p.alpha2 <- ggplot(richness.crab, aes(x=as.factor(stomach_fullness),y=species.richness)) +
  geom_boxplot() +
  labs(y="N. Species",x="Stomach Fullness") +
  scale_y_continuous(limits=c(0,max(richness.crab$species.richness)+2), expand=c(0,0)) +
  theme_bw()
p.alpha2
```
<br>

Save plot.
```{r echo=TRUE}
png(here('figs','fig1b_alpha-div-fullness.png'), res=200, height=600,width=800)
p.alpha2
dev.off()
```
<br>



# nMDS

Run on a presence/absence matrix, using the Jaccard dissimilarity index.
```{r}
crab.sp.mat <- dat %>%
  dplyr::select(crab_id, taxon) %>% distinct() %>%
  mutate(presence=1) %>%
  pivot_wider(id_cols="crab_id", names_from="taxon",values_from="presence", values_fill=0)
crabs <- crab.sp.mat$crab_id
crab.sp.mat <- as.matrix(crab.sp.mat %>% dplyr::select(-crab_id))
rownames(crab.sp.mat) <- crabs
```
```{r eval=FALSE, echo=TRUE}
z <- metaMDS(comm=crab.sp.mat, autotransform=FALSE,distance="jaccard",k=2,maxit=1000,try=40,trymax=100)
```
```{r include=FALSE}
z <- metaMDS(comm=crab.sp.mat, autotransform=FALSE,distance="jaccard",k=2,maxit=1000,try=40,trymax=100)
```
```{r}
message('stress: ', round(z$stress,3))
```
<br>

```{r fig.height=5, fig.width=8}
#extract NMDS scores (x and y coordinates)
z.scores = as.data.frame(scores(z)$sites)
#add columns to data frame 
z.scores$Sample = rownames(z.scores)
z.scores <- left_join(z.scores,dat %>% dplyr::select(crab_id,Site),by=c("Sample"="crab_id"))
z.scores <- z.scores %>% distinct() %>%
  mutate(Site=ifelse(Site=="KAY","Kayak Point 1",ifelse(Site=="KAY-shell", "Kayak Point 2",
                                                        ifelse(Site=="LAR","Larrabee",
                                                               ifelse(Site=="MARPT","March Point",NA)))))

#make hulls
hull.data <- bind_rows((a <- z.scores[z.scores$Site == "Kayak Point 1", ][chull(z.scores[z.scores$Site == 
    "Kayak Point 1", c("NMDS1", "NMDS2")]), ]),
    (b <- z.scores[z.scores$Site == "Kayak Point 2", ][chull(z.scores[z.scores$Site == 
    "Kayak Point 2", c("NMDS1", "NMDS2")]), ]),  # hull values for grp B
    (c <- z.scores[z.scores$Site == "Larrabee", ][chull(z.scores[z.scores$Site == 
    "Larrabee", c("NMDS1", "NMDS2")]), ]),
    (d <- z.scores[z.scores$Site == "March Point", ][chull(z.scores[z.scores$Site == 
    "March Point", c("NMDS1", "NMDS2")]), ]))

#plot
my.nmds <- ggplot(z.scores, aes(x = NMDS1, y = NMDS2)) + 
  geom_point(size = 2, aes(col = Site))+ 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,fill=Site,group=Site),alpha=0.10) + # add the convex hulls
  labs(x = "NMDS1", colour = "Site", y = "NMDS2")  + 
  scale_color_manual(values=c("cyan4","darkblue","darkorange","coral3")) +
  scale_fill_manual(values=c("cyan4","darkblue","darkorange","coral3")) +
  theme_bw() + theme(legend.text=element_text(size=12),
                     legend.title=element_blank(),
                     axis.text=element_blank())
my.nmds
```
<br>

<br>

Save plot.
```{r echo=TRUE}
png(here('figs','pcrg_nmds.png'), res=200, height=800,width=1200)
my.nmds
dev.off()
```
<br>
<br>


# Prey identity

### frequency of all prey groups

### common prey across all sites

Note that I removed March Point data, which comes from only two crab. With the March Point data, there was only one common species across all sites. 

```{r message=FALSE}
# get total crab per site & sites to include in data set.
crab.per.site <- dat %>% group_by(Site) %>% summarise(total_crab=length(unique(crab_id))) %>% ungroup()
my.sites <- crab.per.site$Site[which(crab.per.site$Site != "MARPT")]

# summarise proportion crab per prey ID
crab.per.prey <- dat %>% group_by(Site,taxon,phylum,class,order) %>% 
  summarise(ncrab=length(unique(crab_id))) %>%
  left_join(crab.per.site, by="Site") %>%
  mutate(pcrab=ncrab/total_crab)

ubiq.sp <- dat %>% dplyr::select(Site,taxon) %>%
  distinct() %>%
  mutate(presence=1) %>%
  pivot_wider(names_from=Site,values_from=presence, values_fill=0) %>%
  mutate(total_sites=rowSums(across(c(seq(2,(length(my.sites)+1)))))) %>%
  filter(total_sites>=length(my.sites)) %>%
  left_join(crab.per.prey, by="taxon")
```

```{r}
tmplot <- ggplot(ubiq.sp %>% filter(Site %in% my.sites) %>%
  mutate(Site=ifelse(Site=="KAY","Kayak Point 1",ifelse(Site=="KAY-shell", "Kayak Point 2",
                                                        ifelse(Site=="LAR","Larrabee",
                                                               ifelse(Site=="MARPT","March Point",NA))))) %>%
    mutate(taxon=ifelse(taxon=="Nannochloris sp. MBTD-CMFRI-S048","Nannochloris sp.",taxon)) %>%
             filter(taxon != "Micromonas pusilla"), aes(x=pcrab,y=taxon,fill=Site)) +
  geom_col(position="dodge") +
  labs(x="Prop Crab at Site",y="",subtitle="Species present at all sites") +
  scale_fill_manual(values=c("cyan4","darkblue","darkorange")) + theme_bw() + theme(axis.text.y=element_text(size=11),
                                                                                legend.text=element_text(size=12),
                                                                                legend.title=element_blank())

tmplot
png(here('figs','pcrg_ubiq_sp.png'), res=200, height=800,width=1300)
tmplot
dev.off()

```
