---
title: "Analysis for WSN abstract"
author: "M Fisher"
date: '2022-10-11'
output: 
  html_document:
    toc: yes
    toc_float: yes
---

# Description

Summarize the following from the sequencing run:

1. Prey diversity:
  a. alpha diversity - what proportion of reads were identified down to genus or species? 
  b. beta diversity - What was the unique number of taxa identified per site/month? per site?
  c. nMDS
  d. 
  
2. Prey identification:
  a. what proportion of crab from each site had eaten a given phylum / class?


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(vegan)
library(magrittr)

# User inputs
indir   <- 'data/blast'
run.num <- 1
marker  <- 'lerayXT'
```
```{r data, include=FALSE}
dat <- read_csv(here(indir, 'lerayXT_r1_sample_taxonomy_filtered.csv'))

dat %<>% mutate(Site=ifelse(Site=="KAY" & Sample_num > 18, "KAY-shell",Site))
```
<br>

For this analysis, I removed any identifications that with too little taxonomic depth (information at family or below is missing). What are the sample sizes for this data set?
```{r}
dat.filter <- dat %>%
  filter (rank %in% c("family","species","genus")) %>%
  filter ( !(is.na(class) & is.na(order))) %>%
  filter (species!="Homo sapiens" | is.na(species))

dat.filter %>%
  group_by(Site) %>% summarise(`Crab count`=length(unique(crab_id)))
```
<br>

And let's check the number of crab / site with 1,2 and 3 replicates retained.
```{r}
tmpdat <- dat.filter %>%
  group_by(Site, crab_id) %>% summarise(nreps=length(unique(Replicate)),.groups = "drop")
with(tmpdat, table(Site,nreps))
```
<br>
Good, most crab have all three replicates retained after filtering.

<br>
<br>

# Prey Diversity

### Alpha div

What is the species richness ($\alpha$ diversity) at each site, for **species?** Richness per crab is simply the species richness / total crab from site.
```{r}
prey.freq.site <- dat.filter %>%
  filter(rank=="species") %>%
  group_by(Site) %>%
  summarise(ncrab=length(unique(crab_id)),
            richness=length(unique(species)),
            richness.per.crab=richness/ncrab)
length(dat.filter %>%
  filter(rank=="species") %>% dplyr::select(species) %>% distinct() %>% pull(species))
prey.freq.site
```

<br>


What is the $\alpha$ diversity at each site, inclusive of identifications that only go to genus? I've removed any genus-level identifications that are within the same genus as an existing species-level identification. For example, if I have one ID for *Botrylloides violaceus* then I won't keep a second ID for *Botrylloides*. 
```{r message=FALSE}
# get list of all genuses (geni?) at each site that have species-level IDs
genusOFspecies <- dat.filter %>%
  filter(rank=="species") %>%
  dplyr::select(Site,genus) %>% distinct() %>%
  filter(!is.na(genus)) %>% mutate(genus.exists=1)

removedat <- dat.filter %>%
  left_join(genusOFspecies,by=c("Site","genus")) %>%
  filter(genus.exists==1) %>%
  filter(rank=="genus") %>% dplyr::select(-genus.exists)
  
dat.filtertax <- anti_join(dat.filter,removedat)
  

length(dat.filtertax %>%
  filter(rank=="species" | rank=="genus") %>% dplyr::select(taxon) %>% distinct() %>% pull(taxon))

dat.filtertax %>%
  group_by(Site) %>%
  summarise(ncrab=length(unique(crab_id)),
            richness=length(unique(taxon)),
            richness.per.crab=richness/ncrab)
```

<br>

### Beta div

What is the species richness ($\beta$ diversity) for species? (the ratio between regional and local species diversity)
Note that for this sequencing run only, I include *Hemigrapsus sp* and *Mya sp* to species level. 
```{r}
prey.sp.site.mat <- dat.filter %>%
  filter(rank=="species") %>%
  dplyr::select(Site, taxon) %>% distinct() %>%
  mutate(presence=1) %>%
  pivot_wider(id_cols="Site", names_from="taxon",values_from="presence", values_fill=0)

betadiver(prey.sp.site.mat, method="w")
prey.sp.site.mat$Site

```
The greater the similarity in community composition between multiple communities, the lower the value of β-diversity for that set of communities.

Beta diversity is greater between distant sites (Kayak Point v. Clayton Beach v. March Point) than it is between neighboring sites (Kayak Point eelgrass v. shell).

<br>


### nMDS

```{r}
crab.sp.mat <- dat.filter %>%
  filter(rank=="species") %>%
  dplyr::select(crab_id, taxon) %>% distinct() %>%
  mutate(presence=1) %>%
  pivot_wider(id_cols="crab_id", names_from="taxon",values_from="presence", values_fill=0)
crabs <- crab.sp.mat$crab_id
crab.sp.mat <- as.matrix(crab.sp.mat %>% dplyr::select(-crab_id))
rownames(crab.sp.mat) <- crabs

z <- metaMDS(comm=crab.sp.mat, autotransform=FALSE,distance="bray",k=2,maxit=1000,try=40,trymax=100)
```
<br>

```{r}
stressplot(z)
```
<br>

Disclaimer that **the stress is at 0.2, which is potentially misleading**.


```{r}
#extract NMDS scores (x and y coordinates)
z.scores = as.data.frame(scores(z)$sites)
#add columns to data frame 
z.scores$Sample = rownames(z.scores)
z.scores <- left_join(z.scores,dat.filter %>% dplyr::select(crab_id,Site),by=c("Sample"="crab_id"))
z.scores <- z.scores %>% distinct()
```
```{r fig.height=5, fig.width=8}
ggplot(z.scores, aes(x = NMDS1, y = NMDS2)) + 
    geom_point(size = 4, aes(col = Site))+ 
  geom_text_repel(aes(label=Sample), size=3, col="black", max.overlaps=20) +
    labs(x = "NMDS1", colour = "Group", y = "NMDS2", shape = "Type")  + 
  theme_bw()
```

lol what a mess.
<br>


## Saturation

### Overall
How many unique taxa are added per new crab?
```{r fig.height=3, fig.width=4}
all_sp <- c()
added_sp <- data.frame(ncrab=as.numeric(),
                       nsp=as.numeric())
for(i in seq(1,length(unique(dat.filter$crab_id)))){
  new_sp <- dat.filter %>%
    filter(crab_id==unique(dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,species) %>%
    distinct() %>%
    filter(!(species %in% all_sp)) %>%
    pull(species)
    all_sp <- c(all_sp, new_sp)
  added_sp <- added_sp %>% bind_rows(data.frame(ncrab=i, nsp=length(all_sp)))
}

ggplot(added_sp,aes(x=ncrab,y=nsp)) +
  geom_point() + 
  labs(y="total prey species",x="number of crab") + theme_classic()
```
<br>

what if we look at just Clayton Beach or Kayak Point?
```{r fig.height=3, fig.width=4}
claydat <- filter(dat.filter, Site=="CLAY")
all_sp <- c()
added_sp <- data.frame(site=as.character(),
                       ncrab=as.numeric(),
                       nsp=as.numeric())
for(i in seq(1,length(unique(claydat$crab_id)))){
  new_sp <- claydat %>%
    filter(crab_id==unique(claydat$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,species) %>%
    distinct() %>%
    filter(!(species %in% all_sp)) %>%
    pull(species)
    all_sp <- c(all_sp, new_sp)
  added_sp <- added_sp %>% bind_rows(data.frame(ncrab=i, nsp=length(all_sp)))
}

ggplot(added_sp,aes(x=ncrab,y=nsp)) +
  geom_point() + 
  labs(y="total prey species",x="number of crab") + ggtitle("Clayton Beach") + theme_classic()
```
<br>

At higher taxonomic levels:
```{r fig.height=3, fig.width=4}
## family
all_fam <- c()
added_fam <- data.frame(ncrab=as.numeric(),
                       nfam=as.numeric())
for(i in seq(1,length(unique(dat.filter$crab_id)))){
  new_fam <- dat.filter %>%
    filter(crab_id==unique(dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,family) %>%
    distinct() %>%
    filter(!(family %in% all_fam)) %>%
    pull(family)
    all_fam <- c(all_fam, new_fam)
  added_fam <- added_fam %>% bind_rows(data.frame(ncrab=i, nfam=length(all_fam)))
}
```

```{r}
## order
all_ord <- c()
added_ord <- data.frame(ncrab=as.numeric(),
                       nord=as.numeric())
for(i in seq(1,length(unique(dat.filter$crab_id)))){
  new_ord <- dat.filter %>%
    filter(crab_id==unique(dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,order) %>%
    distinct() %>%
    filter(!(order %in% all_ord)) %>%
    pull(order)
    all_ord <- c(all_ord, new_ord)
  added_ord <- added_ord %>% bind_rows(data.frame(ncrab=i, nord=length(all_ord)))
}
```
```{r}
## class
all_class <- c()
added_class <- data.frame(ncrab=as.numeric(),
                       nclass=as.numeric())
for(i in seq(1,length(unique(dat.filter$crab_id)))){
  new_class <- dat.filter %>%
    filter(crab_id==unique(dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,class) %>%
    distinct() %>%
    filter(!(class %in% all_class)) %>%
    pull(class)
    all_class <- c(all_class, new_class)
  added_class <- added_class %>% bind_rows(data.frame(ncrab=i, nclass=length(all_class)))
}
```
```{r}
## phylum
all_phylum <- c()
added_phylum <- data.frame(ncrab=as.numeric(),
                       nphylum=as.numeric())
for(i in seq(1,length(unique(dat.filter$crab_id)))){
  new_phylum <- dat.filter %>%
    filter(crab_id==unique(dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,phylum) %>%
    distinct() %>%
    filter(!(phylum %in% all_phylum)) %>%
    pull(phylum)
    all_phylum <- c(all_phylum, new_phylum)
  added_phylum <- added_phylum %>% bind_rows(data.frame(ncrab=i, nphylum=length(all_phylum)))
}
```
```{r fig.height=5, fig.width=10}
p1 <- ggplot(added_fam,aes(x=ncrab,y=nfam)) +
  geom_point() + 
  labs(y="total families",x="number of crab") + theme_classic()
p2 <- ggplot(added_ord,aes(x=ncrab,y=nord)) +
  geom_point() + 
  labs(y="total orders",x="number of crab") + theme_classic()
p3 <- ggplot(added_class,aes(x=ncrab,y=nclass)) +
  geom_point() + 
  labs(y="total classes",x="number of crab") + theme_classic()
p4 <- ggplot(added_phylum,aes(x=ncrab,y=nphylum)) +
  geom_point() + 
  labs(y="total phyla",x="number of crab") + theme_classic()
cowplot::plot_grid(p1,p2,p3,p4,ncol=4)
```

<br>

### By site

```{r fig.height=4, fig.width=5}
clay.dat.filter <- dat.filter %>% filter(Site=="CLAY")
## class
all_class <- c()
added_class <- data.frame(ncrab=as.numeric(),
                       nclass=as.numeric())
for(i in seq(1,length(unique(clay.dat.filter$crab_id)))){
  new_class <- clay.dat.filter %>%
    filter(crab_id==unique(clay.dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,class) %>%
    distinct() %>%
    filter(!(class %in% all_class)) %>%
    pull(class)
    all_class <- c(all_class, new_class)
  added_class <- added_class %>% bind_rows(data.frame(ncrab=i, nclass=length(all_class)))
}

ggplot(added_class,aes(x=ncrab,y=nclass)) +
  geom_point() + 
  ylim(c(0,max(added_class$nclass)+2)) +
  labs(y="total classes",x="number of crab") + ggtitle("Clayton Beach") + theme_classic()
```
```{r fig.height=4, fig.width=5}
## order
all_ord <- c()
added_ord <- data.frame(ncrab=as.numeric(),
                       nord=as.numeric())
for(i in seq(1,length(unique(clay.dat.filter$crab_id)))){
  new_ord <- clay.dat.filter %>%
    filter(crab_id==unique(clay.dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,order) %>%
    distinct() %>%
    filter(!(order %in% all_ord)) %>%
    pull(order)
    all_ord <- c(all_ord, new_ord)
  added_ord <- added_ord %>% bind_rows(data.frame(ncrab=i, nord=length(all_ord)))
}

ggplot(added_ord,aes(x=ncrab,y=nord)) +
  geom_point() + 
  ylim(c(0,max(added_ord$nord)+2)) +
  labs(y="total orders",x="number of crab") + ggtitle("Clayton Beach") + theme_classic()
```
<br>

```{r}
kay.dat.filter <- dat.filter %>% filter(Site=="KAY")
## class
all_class <- c()
added_class <- data.frame(ncrab=as.numeric(),
                       nclass=as.numeric())
for(i in seq(1,length(unique(kay.dat.filter$crab_id)))){
  new_class <- kay.dat.filter %>%
    filter(crab_id==unique(kay.dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,class) %>%
    distinct() %>%
    filter(!(class %in% all_class)) %>%
    pull(class)
    all_class <- c(all_class, new_class)
  added_class <- added_class %>% bind_rows(data.frame(ncrab=i, nclass=length(all_class)))
}

kp1 <- ggplot(added_class,aes(x=ncrab,y=nclass)) +
  geom_point() + 
  ylim(c(0,max(added_class$nclass)+2)) +
  labs(y="total classes",x="number of crab") + ggtitle("Kayak Pt - Eelgrass") + theme_classic()
```
```{r}
kay.dat.filter <- dat.filter %>% filter(Site=="KAY-shell")
## class
all_class <- c()
added_class <- data.frame(ncrab=as.numeric(),
                       nclass=as.numeric())
for(i in seq(1,length(unique(kay.dat.filter$crab_id)))){
  new_class <- kay.dat.filter %>%
    filter(crab_id==unique(kay.dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,class) %>%
    distinct() %>%
    filter(!(class %in% all_class)) %>%
    pull(class)
    all_class <- c(all_class, new_class)
  added_class <- added_class %>% bind_rows(data.frame(ncrab=i, nclass=length(all_class)))
}

kp2 <- ggplot(added_class,aes(x=ncrab,y=nclass)) +
  geom_point() + 
  ylim(c(0,max(added_class$nclass)+2)) +
  labs(y="total classes",x="number of crab") + ggtitle("Kayak Pt - Shell") + theme_classic()
```
```{r}
kay.dat.filter <- dat.filter %>% filter(Site=="KAY-shell" | Site=="KAY")
## class
all_class <- c()
added_class <- data.frame(ncrab=as.numeric(),
                       nclass=as.numeric())
for(i in seq(1,length(unique(kay.dat.filter$crab_id)))){
  new_class <- kay.dat.filter %>%
    filter(crab_id==unique(kay.dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,class) %>%
    distinct() %>%
    filter(!(class %in% all_class)) %>%
    pull(class)
    all_class <- c(all_class, new_class)
  added_class <- added_class %>% bind_rows(data.frame(ncrab=i, nclass=length(all_class)))
}

kp3 <- ggplot(added_class,aes(x=ncrab,y=nclass)) +
  geom_point() + 
  ylim(c(0,max(added_class$nclass)+2)) +
  labs(y="total classes",x="number of crab") + ggtitle("Kayak Pt - All") + theme_classic()
```
````{r fig.height=3, fig.width=8}
plot_grid(kp1,kp2,kp3,ncol=3)
```
<br>


```{r}
kay.dat.filter <- dat.filter %>% filter(Site=="KAY")
## order
all_order <- c()
added_order <- data.frame(ncrab=as.numeric(),
                       norder=as.numeric())
for(i in seq(1,length(unique(kay.dat.filter$crab_id)))){
  new_order <- kay.dat.filter %>%
    filter(crab_id==unique(kay.dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,order) %>%
    distinct() %>%
    filter(!(order %in% all_order)) %>%
    pull(order)
    all_order <- c(all_order, new_order)
  added_order <- added_order %>% bind_rows(data.frame(ncrab=i, norder=length(all_order)))
}

kp1 <- ggplot(added_order,aes(x=ncrab,y=norder)) +
  geom_point() + 
  ylim(c(0,max(added_order$norder)+2)) +
  labs(y="total orders",x="number of crab") + ggtitle("Kayak Pt - Eelgrass") + theme_classic()
```
```{r}
kay.dat.filter <- dat.filter %>% filter(Site=="KAY-shell")
## order
all_order <- c()
added_order <- data.frame(ncrab=as.numeric(),
                       norder=as.numeric())
for(i in seq(1,length(unique(kay.dat.filter$crab_id)))){
  new_order <- kay.dat.filter %>%
    filter(crab_id==unique(kay.dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,order) %>%
    distinct() %>%
    filter(!(order %in% all_order)) %>%
    pull(order)
    all_order <- c(all_order, new_order)
  added_order <- added_order %>% bind_rows(data.frame(ncrab=i, norder=length(all_order)))
}

kp2 <- ggplot(added_order,aes(x=ncrab,y=norder)) +
  geom_point() + 
  ylim(c(0,max(added_order$norder)+2)) +
  labs(y="total orderes",x="number of crab") + ggtitle("Kayak Pt - Shell") + theme_classic()
```
```{r}
kay.dat.filter <- dat.filter %>% filter(Site=="KAY-shell" | Site=="KAY")
## order
all_order <- c()
added_order <- data.frame(ncrab=as.numeric(),
                       norder=as.numeric())
for(i in seq(1,length(unique(kay.dat.filter$crab_id)))){
  new_order <- kay.dat.filter %>%
    filter(crab_id==unique(kay.dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,order) %>%
    distinct() %>%
    filter(!(order %in% all_order)) %>%
    pull(order)
    all_order <- c(all_order, new_order)
  added_order <- added_order %>% bind_rows(data.frame(ncrab=i, norder=length(all_order)))
}

kp3 <- ggplot(added_order,aes(x=ncrab,y=norder)) +
  geom_point() + 
  ylim(c(0,max(added_order$norder)+2)) +
  labs(y="total orders",x="number of crab") + ggtitle("Kayak Pt - All") + theme_classic()
```
````{r fig.height=3, fig.width=8}
plot_grid(kp1,kp2,kp3,ncol=3)
```

<br>
<br>


# Prey Identification

## Overall

```{r}
taxa_summary <- dat.filter %>%
  mutate(total_reads=sum(dat.filter$nReads),
         total_crab=length(unique((dat.filter$crab_id)))) %>%
  group_by(taxon, rank,class,phylum) %>% 
  summarise(n_crab=length(unique(crab_id)),
            prop_crab=n_crab/total_crab,
            n_reads=sum(nReads),
            prop_reads=n_reads/total_reads) %>%
  distinct() %>% arrange(n_crab)
```
I've removed the species that only showed up in one crab. That's a total of `r length(unique(taxa_summary %>% filter(n_crab==1) %>% pull(taxon)))` unique taxa.

```{r fig.height=8,fig.width=9}
plotdat <- taxa_summary %>%
  filter(n_crab > 1) %>%
  dplyr::select(-prop_crab,-n_reads) %>%
  arrange(n_crab) %>%    # First sort by val. This sort the dataframe but NOT the factor levels
  mutate(taxon=factor(taxon, levels=taxon))
plotdat %>%
  pivot_longer(cols=c(n_crab,prop_reads),names_to="metric",values_to="value") %>%
  mutate(metric=ifelse(metric=="n_crab","Crab Count","Prop Sequence Reads")) %>%
  ggplot(aes(y=value,x=factor(taxon, levels=plotdat$taxon))) +
  facet_grid(cols=vars(metric),scales="free") +
  geom_col(aes(fill=rank)) +
  coord_flip() + theme_bw() + theme(axis.title=element_blank(),
                                    strip.text=element_text(size=12),
                                    axis.text.x=element_text(size=12),
                                    axis.text.y=element_text(size=10))
```
```{r fig.height=5, fig.width=9}
plot1 <- dat.filter %>%
  group_by(phylum) %>%
  summarise(n_crab=length(unique(crab_id)),
            n_taxon=length(unique(taxon))) %>%
  ggplot(aes(y=n_taxon,x=n_crab, col=phylum)) +
  geom_point() +
  geom_text_repel(aes(label=phylum), size=3, col="black", max.overlaps=30) +
  labs(x="Number of crab",y="Number unique taxa") +
  theme_bw() + theme(axis.text.x=element_text(size=10),
                                    axis.text.y=element_text(size=10),
                     legend.position="none")
plot2 <- dat.filter %>%
  filter(rank=="species") %>%
  group_by(phylum) %>%
  summarise(n_crab=length(unique(crab_id)),
            n_taxon=length(unique(taxon))) %>%
  ggplot(aes(y=n_taxon,x=n_crab, col=phylum)) +
  geom_point() +
  geom_text_repel(aes(label=phylum), size=3, col="black", max.overlaps=30) +
  labs(x="Number of crab",y="Number unique species") +
  theme_bw() + theme(axis.text.x=element_text(size=10),
                                    axis.text.y=element_text(size=10),
                     legend.position="none")
plot_grid(plot1,plot2,ncol=2)
```
<br>


save this summary
```{r eval=FALSE, echo=TRUE}
write_csv(taxa_summary,here(indir, 'lerayXT_r1_taxonomy_summary.csv'))
```
<br>
<br>

## By site
```{r}
site.totals <- dat.filter %>% group_by(Site) %>% 
  summarise(total_reads=sum(nReads),
         total_crab=length(unique(crab_id)))
taxa_summary_bysite <- dat.filter %>%
  group_by(taxon, rank, Site) %>% 
  summarise(n_crab=length(unique(crab_id)),
            n_reads=sum(nReads)) %>%
  left_join(site.totals,by="Site") %>%
  mutate(prop_crab=n_crab/total_crab,
         prop_reads=n_reads/total_reads) %>%
  distinct() %>% arrange(n_crab)
```
```{r fig.height=7,fig.width=9}
plotdat <- taxa_summary_bysite %>%
  filter(n_crab > 1) %>%
  dplyr::select(-n_crab,-n_reads) %>%
  arrange(prop_crab) %>%    # First sort by val. This sort the dataframe but NOT the factor levels
  mutate(taxon=factor(taxon, levels=unique(taxon)))
plotdat %>%
  ggplot(aes(y=prop_crab,x=factor(taxon, levels=unique(plotdat$taxon)))) +
  facet_grid(cols=vars(Site),scales="free") +
  geom_col(aes(fill=rank)) +
  labs(x="Prop Crab per Site") +
  ylim(c(0,1)) +
  coord_flip() + theme_bw() + theme(axis.title=element_blank(),
                                    strip.text=element_text(size=12),
                                    axis.text.x=element_text(size=12),
                                    axis.text.y=element_text(size=10))
```
<br>

save this summary
```{r eval=FALSE, echo=TRUE}
write_csv(taxa_summary_bysite,here(indir, 'lerayXT_r1_taxonomy_summary_by_site.csv'))
```
<br>

What about at higher orders of taxonomy? For only those identifications to species.

```{r}
site.totals.sp <- dat.filter %>% filter(rank=="species") %>%
  group_by(Site) %>% 
  summarise(total_reads=sum(nReads),
         total_crab=length(unique(crab_id)))
class_summary_bysite <- dat.filter %>%
  filter(rank=="species") %>%
  group_by(class, Site) %>% 
  summarise(n_crab=length(unique(crab_id)),
            n_reads=sum(nReads)) %>%
  left_join(site.totals.sp,by="Site") %>%
  mutate(prop_crab=n_crab/total_crab,
         prop_reads=n_reads/total_reads) %>%
  distinct() %>% arrange(n_crab)
```
```{r fig.height=7,fig.width=8}
plotdat <- class_summary_bysite %>%
  dplyr::select(class,Site,n_crab,prop_reads) %>%
  arrange(n_crab) %>%    # First sort by val. This sort the dataframe but NOT the factor levels
  mutate(class=factor(class, levels=unique(class)))
plotdat %>%
  pivot_longer(cols=c(n_crab,prop_reads),names_to="metric",values_to="value") %>%
  mutate(metric=ifelse(metric=="n_crab","Crab Count","Prop Sequence Reads")) %>%
  ggplot(aes(y=value,x=factor(class, levels=unique(plotdat$class)))) +
  facet_grid(cols=vars(metric),scales="free") +
  geom_col(aes(fill=Site), position="dodge") +
  ggtitle("Classes") +
  coord_flip() + theme_bw() + theme(axis.title=element_blank(),
                                    strip.text=element_text(size=12),
                                    axis.text.x=element_text(size=12),
                                    axis.text.y=element_text(size=10))
```
<br>
```{r}
phylum_summary_bysite <- dat.filter %>%
  filter(rank=="species") %>%
  group_by(phylum, Site) %>% 
  summarise(n_crab=length(unique(crab_id)),
            n_reads=sum(nReads)) %>%
  left_join(site.totals.sp,by="Site") %>%
  mutate(prop_crab=n_crab/total_crab,
         prop_reads=n_reads/total_reads) %>%
  distinct() %>% arrange(n_crab)
```
```{r fig.height=5,fig.width=7}
plotdat <- phylum_summary_bysite %>%
  dplyr::select(phylum,Site,n_crab,prop_reads) %>%
  arrange(n_crab) %>%    # First sort by val. This sort the dataframe but NOT the factor levels
  mutate(class=factor(phylum, levels=unique(phylum)))
plotdat %>%
  pivot_longer(cols=c(n_crab,prop_reads),names_to="metric",values_to="value") %>%
  mutate(metric=ifelse(metric=="n_crab","Crab Count","Prop Sequence Reads")) %>%
  ggplot(aes(y=value,x=factor(phylum, levels=unique(plotdat$class)))) +
  facet_grid(cols=vars(metric),scales="free") +
  geom_col(aes(fill=Site), position="dodge") +
  ggtitle("Phyla") +
  coord_flip() + theme_bw() + theme(axis.title=element_blank(),
                                    strip.text=element_text(size=12),
                                    axis.text.x=element_text(size=12),
                                    axis.text.y=element_text(size=10))
```
<br>

What's going on with the "NA" phylum?
```{r}
unique(dat.filter %>% filter(is.na(phylum)) %>% pull(taxon))
```

Some larger algae, some single-celled algae.



