---
title: "Analysis for WSN abstract"
author: "M Fisher"
date: '2022-10-11'
output: 
  html_document:
    toc: yes
    toc_float: yes
---

# Description

Summarize the following from the sequencing run:

1. Prey diversity:
  a. alpha diversity - what proportion of reads were identified down to genus or species? overall and per crab
  b. beta diversity - What was the unique number of taxa identified per site/month? per site?
  c. taxonomic distinctness (vegan doc)[https://search.r-project.org/CRAN/refmans/vegan/html/taxondive.html]
  d. nMDS
  
2. Prey identification:
  a. what proportion of crab from each site had eaten a given phylum / class?
  b. what proportion of crab from each site had eaten a given group of species, based on common names?


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(vegan)
library(magrittr)

# User inputs
indir   <- 'data/blast'
run.num <- 1
marker  <- 'lerayXT'
```
```{r data, include=FALSE}
dat <- read_csv(here(indir, 'lerayXT_r1_sample_taxonomy_filtered.csv'))

dat %<>% mutate(Site=ifelse(Site=="KAY" & Sample_num > 18, "KAY-shell",Site))
```
<br>

For this analysis, I removed any identifications that with too little taxonomic depth (information at family or below is missing.
```{r}
dat.filter <- dat %>%
  filter (rank %in% c("family","species","genus"))
```


What are the final sample sizes for this data set?
```{r}
dat.filter %>%
  group_by(Site) %>% summarise(`Crab count`=length(unique(crab_id)))
```
<br>

And let's check the number of crab / site with 1,2 and 3 replicates retained.
```{r}
tmpdat <- dat.filter %>%
  group_by(Site, crab_id) %>% summarise(nreps=length(unique(Replicate)),.groups = "drop")
with(tmpdat, table(Site,nreps))
```
<br>
Good, most crab have all three replicates retained after filtering.

I'm interested in two versions of the final filtered data set: one including only species-level taxonomy (**species**), and one including taxa that are unique to a given site (**unique taxa**). For example, if a taxon is only identified down to genus, and that genus is found in another ID at the same site, it will be removed; but if it is not found in another ID at the same site, it is retained. I'll save this as a separate dataset.
```{r}
species_gen <- dat.filter %>% 
  filter(rank=="species") %>%
  dplyr::select(Site, genus) %>%
  distinct() %>%
  mutate(genus_in_speciesdat=1)

species_fam <- dat.filter %>% 
  filter(rank=="species") %>%
  dplyr::select(Site, family) %>%
  distinct() %>%
  mutate(family_in_speciesdat=1)

taxa.dat <- dat.filter %>%
  left_join(species_gen,by=c("Site","genus")) %>%
  left_join(species_fam,by=c("Site","family")) %>%
  filter(rank=="species" | (rank=="genus" & is.na(genus_in_speciesdat)) | (rank=="family" & is.na(family_in_speciesdat))) %>%
  dplyr::select(-genus_in_speciesdat,-family_in_speciesdat)
```
```{r echo=TRUE, eval=FALSE}
write_csv(taxa.dat, (here(indir, 'lerayXT_r1_sample_taxonomy_filtered_SiteUniqueTaxa.csv')))
```
<br>

And I'll save a list of the species / taxa to look up common names:
```{r}
taxa_list <- taxa.dat %>% dplyr::select(rank,taxon,phylum,class,order,family,genus,species) %>% distinct()
species_list <- dat.filter %>% filter(rank=="species") %>%
  dplyr::select(Site,crab_id,taxon,rank,phylum,class,order,family,genus,species) %>% 
  group_by(taxon,rank,phylum,class,order,family,genus,species) %>%
  summarise(n_sites=length(unique(Site)),
         n_crab=length(unique(crab_id))) %>%
  arrange(n_crab)
```
```{r echo=TRUE}
write_csv(taxa_list, here('data','results','lerayXT_r1_taxa_common_names.csv'))
write_csv(species_list, here('data','results','lerayXT_r1_species_common_names.csv'))
```
<br>
<br>

# Prey Diversity

### Alpha div

What is the species richness ($\alpha$ diversity) at each site, for **species?** Richness per crab is simply the species richness / total crab from site.
```{r}
prey.freq.site <- dat.filter %>%
  filter(rank=="species") %>%
  group_by(Site) %>%
  summarise(ncrab=length(unique(crab_id)),
            species.richness=length(unique(species)),
            mean.richness.per.crab=species.richness/ncrab)
length(dat.filter %>%
  filter(rank=="species") %>% dplyr::select(species) %>% distinct() %>% pull(species))
prey.freq.site
```

<br>


What is the $\alpha$ diversity at each site, for unique taxa?
```{r message=FALSE}
length(taxa.dat %>% dplyr::select(taxon) %>% distinct() %>% pull(taxon))

taxa.dat %>%
  group_by(Site) %>%
  summarise(ncrab=length(unique(crab_id)),
            taxa.richness=length(unique(taxon)),
            mean.richness.per.crab=taxa.richness/ncrab)
```

<br>

### Beta div

What is the $\beta$ diversity for species? (the ratio between regional and local species diversity)
```{r}
prey.sp.site.mat <- dat.filter %>%
  filter(rank=="species") %>%
  dplyr::select(Site, taxon) %>% distinct() %>%
  mutate(presence=1) %>%
  pivot_wider(id_cols="Site", names_from="taxon",values_from="presence", values_fill=0)

betadiver(prey.sp.site.mat, method="w")
prey.sp.site.mat$Site

```
The greater the similarity in community composition between multiple communities, the lower the value of β-diversity for that set of communities.

Beta diversity is greater between distant sites (Kayak Point v. Clayton Beach v. March Point) than it is between neighboring sites (Kayak Point eelgrass v. shell).

<br>

What is the $\beta$ diversity for taxa?
```{r}
taxa.sp.site.mat <- taxa.dat %>%
  dplyr::select(Site, taxon) %>% distinct() %>%
  mutate(presence=1) %>%
  pivot_wider(id_cols="Site", names_from="taxon",values_from="presence", values_fill=0)

betadiver(taxa.sp.site.mat, method="w")
taxa.sp.site.mat$Site

```

<br>
The same pattern holds as above.
<br>


### nMDS

**For Species**
```{r}
crab.sp.mat <- dat.filter %>%
  filter(rank=="species") %>%
  dplyr::select(crab_id, taxon) %>% distinct() %>%
  mutate(presence=1) %>%
  pivot_wider(id_cols="crab_id", names_from="taxon",values_from="presence", values_fill=0)
crabs <- crab.sp.mat$crab_id
crab.sp.mat <- as.matrix(crab.sp.mat %>% dplyr::select(-crab_id))
rownames(crab.sp.mat) <- crabs

z <- metaMDS(comm=crab.sp.mat, autotransform=FALSE,distance="bray",k=2,maxit=1000,try=40,trymax=100)
```
<br>

```{r fig.height=4,fig.width=5}
stressplot(z)
```
<br>

Disclaimer that **the stress is at 0.2, which is potentially misleading**.


```{r}
#extract NMDS scores (x and y coordinates)
z.scores = as.data.frame(scores(z)$sites)
#add columns to data frame 
z.scores$Sample = rownames(z.scores)
z.scores <- left_join(z.scores,dat.filter %>% dplyr::select(crab_id,Site),by=c("Sample"="crab_id"))
z.scores <- z.scores %>% distinct()
```
```{r fig.height=5, fig.width=8}
ggplot(z.scores, aes(x = NMDS1, y = NMDS2)) + 
    geom_point(size = 4, aes(col = Site))+ 
  geom_text_repel(aes(label=Sample), size=3, col="black", max.overlaps=20) +
    labs(x = "NMDS1", colour = "Group", y = "NMDS2", shape = "Type")  + 
  theme_bw()
```

lol what a mess.
<br>
<br>


## Saturation

### Overall
How many unique species are added per new crab?
```{r fig.height=3, fig.width=4}
all_sp <- c()
added_sp <- data.frame(ncrab=as.numeric(),
                       nsp=as.numeric())
for(i in seq(1,length(unique(dat.filter$crab_id)))){
  new_sp <- dat.filter %>%
    filter(crab_id==unique(dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,species) %>%
    distinct() %>%
    filter(!(species %in% all_sp)) %>%
    pull(species)
    all_sp <- c(all_sp, new_sp)
  added_sp <- added_sp %>% bind_rows(data.frame(ncrab=i, nsp=length(all_sp)))
}

ggplot(added_sp,aes(x=ncrab,y=nsp)) +
  geom_point() + 
  labs(y="total prey species",x="number of crab") + theme_classic()
```
<br>

At higher taxonomic levels:
```{r fig.height=3, fig.width=4}
## family
all_fam <- c()
added_fam <- data.frame(ncrab=as.numeric(),
                       nfam=as.numeric())
for(i in seq(1,length(unique(dat.filter$crab_id)))){
  new_fam <- dat.filter %>%
    filter(crab_id==unique(dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,family) %>%
    distinct() %>%
    filter(!(family %in% all_fam)) %>%
    pull(family)
    all_fam <- c(all_fam, new_fam)
  added_fam <- added_fam %>% bind_rows(data.frame(ncrab=i, nfam=length(all_fam)))
}
```

```{r}
## order
all_ord <- c()
added_ord <- data.frame(ncrab=as.numeric(),
                       nord=as.numeric())
for(i in seq(1,length(unique(dat.filter$crab_id)))){
  new_ord <- dat.filter %>%
    filter(crab_id==unique(dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,order) %>%
    distinct() %>%
    filter(!(order %in% all_ord)) %>%
    pull(order)
    all_ord <- c(all_ord, new_ord)
  added_ord <- added_ord %>% bind_rows(data.frame(ncrab=i, nord=length(all_ord)))
}
```
```{r}
## class
all_class <- c()
added_class <- data.frame(ncrab=as.numeric(),
                       nclass=as.numeric())
for(i in seq(1,length(unique(dat.filter$crab_id)))){
  new_class <- dat.filter %>%
    filter(crab_id==unique(dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,class) %>%
    distinct() %>%
    filter(!(class %in% all_class)) %>%
    pull(class)
    all_class <- c(all_class, new_class)
  added_class <- added_class %>% bind_rows(data.frame(ncrab=i, nclass=length(all_class)))
}
```
```{r}
## phylum
all_phylum <- c()
added_phylum <- data.frame(ncrab=as.numeric(),
                       nphylum=as.numeric())
for(i in seq(1,length(unique(dat.filter$crab_id)))){
  new_phylum <- dat.filter %>%
    filter(crab_id==unique(dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,phylum) %>%
    distinct() %>%
    filter(!(phylum %in% all_phylum)) %>%
    pull(phylum)
    all_phylum <- c(all_phylum, new_phylum)
  added_phylum <- added_phylum %>% bind_rows(data.frame(ncrab=i, nphylum=length(all_phylum)))
}
```
```{r fig.height=5, fig.width=10}
p1 <- ggplot(added_fam,aes(x=ncrab,y=nfam)) +
  geom_point() + 
  labs(y="total families",x="number of crab") + theme_classic()
p2 <- ggplot(added_ord,aes(x=ncrab,y=nord)) +
  geom_point() + 
  labs(y="total orders",x="number of crab") + theme_classic()
p3 <- ggplot(added_class,aes(x=ncrab,y=nclass)) +
  geom_point() + 
  labs(y="total classes",x="number of crab") + theme_classic()
p4 <- ggplot(added_phylum,aes(x=ncrab,y=nphylum)) +
  geom_point() + 
  labs(y="total phyla",x="number of crab") + theme_classic()
cowplot::plot_grid(p1,p2,p3,p4,ncol=4)
```
<br>

<br>

### By site

```{r fig.height=3, fig.width=4}
claydat <- filter(dat.filter, Site=="CLAY")
all_sp <- c()
added_sp <- data.frame(site=as.character(),
                       ncrab=as.numeric(),
                       nsp=as.numeric())
for(i in seq(1,length(unique(claydat$crab_id)))){
  new_sp <- claydat %>%
    filter(crab_id==unique(claydat$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,species) %>%
    distinct() %>%
    filter(!(species %in% all_sp)) %>%
    pull(species)
    all_sp <- c(all_sp, new_sp)
  added_sp <- added_sp %>% bind_rows(data.frame(ncrab=i, nsp=length(all_sp)))
}

ggplot(added_sp,aes(x=ncrab,y=nsp)) +
  geom_point() + 
  labs(y="total prey species",x="number of crab") + ggtitle("Clayton Beach") + theme_classic()
```


```{r fig.height=4, fig.width=5}
clay.dat.filter <- dat.filter %>% filter(Site=="CLAY")
## class
all_class <- c()
added_class <- data.frame(ncrab=as.numeric(),
                       nclass=as.numeric())
for(i in seq(1,length(unique(clay.dat.filter$crab_id)))){
  new_class <- clay.dat.filter %>%
    filter(crab_id==unique(clay.dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,class) %>%
    distinct() %>%
    filter(!(class %in% all_class)) %>%
    pull(class)
    all_class <- c(all_class, new_class)
  added_class <- added_class %>% bind_rows(data.frame(ncrab=i, nclass=length(all_class)))
}

ggplot(added_class,aes(x=ncrab,y=nclass)) +
  geom_point() + 
  ylim(c(0,max(added_class$nclass)+2)) +
  labs(y="total classes",x="number of crab") + ggtitle("Clayton Beach") + theme_classic()
```
```{r fig.height=4, fig.width=5}
## order
all_ord <- c()
added_ord <- data.frame(ncrab=as.numeric(),
                       nord=as.numeric())
for(i in seq(1,length(unique(clay.dat.filter$crab_id)))){
  new_ord <- clay.dat.filter %>%
    filter(crab_id==unique(clay.dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,order) %>%
    distinct() %>%
    filter(!(order %in% all_ord)) %>%
    pull(order)
    all_ord <- c(all_ord, new_ord)
  added_ord <- added_ord %>% bind_rows(data.frame(ncrab=i, nord=length(all_ord)))
}

ggplot(added_ord,aes(x=ncrab,y=nord)) +
  geom_point() + 
  ylim(c(0,max(added_ord$nord)+2)) +
  labs(y="total orders",x="number of crab") + ggtitle("Clayton Beach") + theme_classic()
```
<br>

```{r}
kay.dat.filter <- dat.filter %>% filter(Site=="KAY")
## class
all_class <- c()
added_class <- data.frame(ncrab=as.numeric(),
                       nclass=as.numeric())
for(i in seq(1,length(unique(kay.dat.filter$crab_id)))){
  new_class <- kay.dat.filter %>%
    filter(crab_id==unique(kay.dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,class) %>%
    distinct() %>%
    filter(!(class %in% all_class)) %>%
    pull(class)
    all_class <- c(all_class, new_class)
  added_class <- added_class %>% bind_rows(data.frame(ncrab=i, nclass=length(all_class)))
}

kp1 <- ggplot(added_class,aes(x=ncrab,y=nclass)) +
  geom_point() + 
  ylim(c(0,max(added_class$nclass)+2)) +
  labs(y="total classes",x="number of crab") + ggtitle("Kayak Pt - Eelgrass") + theme_classic()
```
```{r}
kay.dat.filter <- dat.filter %>% filter(Site=="KAY-shell")
## class
all_class <- c()
added_class <- data.frame(ncrab=as.numeric(),
                       nclass=as.numeric())
for(i in seq(1,length(unique(kay.dat.filter$crab_id)))){
  new_class <- kay.dat.filter %>%
    filter(crab_id==unique(kay.dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,class) %>%
    distinct() %>%
    filter(!(class %in% all_class)) %>%
    pull(class)
    all_class <- c(all_class, new_class)
  added_class <- added_class %>% bind_rows(data.frame(ncrab=i, nclass=length(all_class)))
}

kp2 <- ggplot(added_class,aes(x=ncrab,y=nclass)) +
  geom_point() + 
  ylim(c(0,max(added_class$nclass)+2)) +
  labs(y="total classes",x="number of crab") + ggtitle("Kayak Pt - Shell") + theme_classic()
```
```{r}
kay.dat.filter <- dat.filter %>% filter(Site=="KAY-shell" | Site=="KAY")
## class
all_class <- c()
added_class <- data.frame(ncrab=as.numeric(),
                       nclass=as.numeric())
for(i in seq(1,length(unique(kay.dat.filter$crab_id)))){
  new_class <- kay.dat.filter %>%
    filter(crab_id==unique(kay.dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,class) %>%
    distinct() %>%
    filter(!(class %in% all_class)) %>%
    pull(class)
    all_class <- c(all_class, new_class)
  added_class <- added_class %>% bind_rows(data.frame(ncrab=i, nclass=length(all_class)))
}

kp3 <- ggplot(added_class,aes(x=ncrab,y=nclass)) +
  geom_point() + 
  ylim(c(0,max(added_class$nclass)+2)) +
  labs(y="total classes",x="number of crab") + ggtitle("Kayak Pt - All") + theme_classic()
```
````{r fig.height=3, fig.width=8}
plot_grid(kp1,kp2,kp3,ncol=3)
```
<br>


```{r}
kay.dat.filter <- dat.filter %>% filter(Site=="KAY")
## order
all_order <- c()
added_order <- data.frame(ncrab=as.numeric(),
                       norder=as.numeric())
for(i in seq(1,length(unique(kay.dat.filter$crab_id)))){
  new_order <- kay.dat.filter %>%
    filter(crab_id==unique(kay.dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,order) %>%
    distinct() %>%
    filter(!(order %in% all_order)) %>%
    pull(order)
    all_order <- c(all_order, new_order)
  added_order <- added_order %>% bind_rows(data.frame(ncrab=i, norder=length(all_order)))
}

kp1 <- ggplot(added_order,aes(x=ncrab,y=norder)) +
  geom_point() + 
  ylim(c(0,max(added_order$norder)+2)) +
  labs(y="total orders",x="number of crab") + ggtitle("Kayak Pt - Eelgrass") + theme_classic()
```
```{r}
kay.dat.filter <- dat.filter %>% filter(Site=="KAY-shell")
## order
all_order <- c()
added_order <- data.frame(ncrab=as.numeric(),
                       norder=as.numeric())
for(i in seq(1,length(unique(kay.dat.filter$crab_id)))){
  new_order <- kay.dat.filter %>%
    filter(crab_id==unique(kay.dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,order) %>%
    distinct() %>%
    filter(!(order %in% all_order)) %>%
    pull(order)
    all_order <- c(all_order, new_order)
  added_order <- added_order %>% bind_rows(data.frame(ncrab=i, norder=length(all_order)))
}

kp2 <- ggplot(added_order,aes(x=ncrab,y=norder)) +
  geom_point() + 
  ylim(c(0,max(added_order$norder)+2)) +
  labs(y="total orderes",x="number of crab") + ggtitle("Kayak Pt - Shell") + theme_classic()
```
```{r}
kay.dat.filter <- dat.filter %>% filter(Site=="KAY-shell" | Site=="KAY")
## order
all_order <- c()
added_order <- data.frame(ncrab=as.numeric(),
                       norder=as.numeric())
for(i in seq(1,length(unique(kay.dat.filter$crab_id)))){
  new_order <- kay.dat.filter %>%
    filter(crab_id==unique(kay.dat.filter$crab_id)[i]) %>%
    filter(rank=="species") %>%
    dplyr::select(crab_id,order) %>%
    distinct() %>%
    filter(!(order %in% all_order)) %>%
    pull(order)
    all_order <- c(all_order, new_order)
  added_order <- added_order %>% bind_rows(data.frame(ncrab=i, norder=length(all_order)))
}

kp3 <- ggplot(added_order,aes(x=ncrab,y=norder)) +
  geom_point() + 
  ylim(c(0,max(added_order$norder)+2)) +
  labs(y="total orders",x="number of crab") + ggtitle("Kayak Pt - All") + theme_classic()
```
````{r fig.height=3, fig.width=8}
plot_grid(kp1,kp2,kp3,ncol=3)
```

<br>
<br>


# Prey Identification

For this section, I'm reading in a species-level data set that has been filtered for random nonspecific species (e.g., "cultured zooplankton") and for non-native species that have no history of invasion in the area.
```{r echo=TRUE}
taxa_summary <- read_csv(here('data','results','lerayXT_r1_species_filtered_common_names.csv'))
```
<br>
<br>

## Overall
```{r message=FALSE}
reads_summary <- dat.filter %>%
  filter(taxon %in% taxa_summary$taxon) %>%
  group_by(rank,taxon) %>%
  summarise(n_reads=sum(nReads))

taxa_summary <- taxa_summary %>%
  left_join(reads_summary, by=c("rank","taxon"))
```
<br>

I've removed the species that only showed up in one crab. That's a total of `r length(unique(taxa_summary %>% filter(n_crab==1) %>% pull(taxon)))` unique taxa.

```{r fig.height=8,fig.width=9}
plotdat <- taxa_summary %>%
  filter(n_crab > 1) %>%
  mutate(`Mean Reads / Crab`=n_reads/n_crab) %>%
  arrange(n_crab) %>%    # First sort by val. This sort the dataframe but NOT the factor levels
  mutate(taxon=factor(taxon, levels=taxon))
plotdat %>%
  pivot_longer(cols=c(n_crab,n_reads, `Mean Reads / Crab`),names_to="metric",values_to="value") %>%
  mutate(metric=ifelse(metric=="n_crab","Crab Count",
                       ifelse(metric=="n_reads","Total Reads", metric))) %>%
  ggplot(aes(y=value,x=factor(taxon, levels=plotdat$taxon))) +
  facet_grid(cols=vars(metric),scales="free") +
  geom_col() +
  coord_flip() + theme_bw() + theme(axis.title=element_blank(),
                                    strip.text=element_text(size=12),
                                    axis.text.x=element_text(size=12),
                                    axis.text.y=element_text(size=10))
```
```{r fig.height=5, fig.width=9}
plot1 <- dat.filter %>%
  group_by(phylum) %>%
  summarise(n_crab=length(unique(crab_id)),
            n_taxon=length(unique(taxon))) %>%
  ggplot(aes(y=n_taxon,x=n_crab, col=phylum)) +
  geom_point() +
  geom_text_repel(aes(label=phylum), size=3, col="black", max.overlaps=30) +
  labs(x="Number of crab",y="Number unique taxa") +
  theme_bw() + theme(axis.text.x=element_text(size=10),
                                    axis.text.y=element_text(size=10),
                     legend.position="none")
plot2 <- taxa_summary %>%
  group_by(phylum) %>%
  summarise(n_crab=sum(n_crab),
            n_taxon=length(unique(taxon))) %>%
  ggplot(aes(y=n_taxon,x=n_crab, col=phylum)) +
  geom_point() +
  geom_text_repel(aes(label=phylum), size=3, col="black", max.overlaps=30) +
  labs(x="Number of crab",y="Number unique (filtered) species") +
  theme_bw() + theme(axis.text.x=element_text(size=10),
                                    axis.text.y=element_text(size=10),
                     legend.position="none")
plot_grid(plot1,plot2,ncol=2)
```
<br>
<br>

## By site
```{r}
site.totals <- dat.filter %>% group_by(Site) %>% 
  summarise(total_reads=sum(nReads),
         total_crab=length(unique(crab_id)))
taxa_summary_bysite <- dat.filter %>%
  group_by(taxon, rank, Site) %>% 
  summarise(n_crab=length(unique(crab_id)),
            n_reads=sum(nReads)) %>%
  left_join(site.totals,by="Site") %>%
  mutate(prop_crab=n_crab/total_crab,
         prop_reads=n_reads/total_reads) %>%
  distinct() %>% arrange(n_crab)
```
```{r fig.height=7,fig.width=9}
plotdat <- taxa_summary_bysite %>%
  filter(n_crab > 1) %>%
  dplyr::select(-n_crab,-n_reads) %>%
  arrange(prop_crab) %>%    # First sort by val. This sort the dataframe but NOT the factor levels
  mutate(taxon=factor(taxon, levels=unique(taxon)))
plotdat %>%
  ggplot(aes(y=prop_crab,x=factor(taxon, levels=unique(plotdat$taxon)))) +
  facet_grid(cols=vars(Site),scales="free") +
  geom_col(aes(fill=rank)) +
  labs(x="Prop Crab per Site") +
  ylim(c(0,1)) +
  coord_flip() + theme_bw() + theme(axis.title=element_blank(),
                                    strip.text=element_text(size=12),
                                    axis.text.x=element_text(size=12),
                                    axis.text.y=element_text(size=10))
```
<br>

save this summary
```{r eval=FALSE, echo=TRUE}
write_csv(taxa_summary_bysite,here(indir, 'lerayXT_r1_taxonomy_summary_by_site.csv'))
```
<br>

What about at higher orders of taxonomy? For only those identifications to species.

```{r}
site.totals.sp <- dat.filter %>% filter(rank=="species") %>%
  group_by(Site) %>% 
  summarise(total_reads=sum(nReads),
         total_crab=length(unique(crab_id)))
class_summary_bysite <- dat.filter %>%
  filter(rank=="species") %>%
  group_by(class, Site) %>% 
  summarise(n_crab=length(unique(crab_id)),
            n_reads=sum(nReads)) %>%
  left_join(site.totals.sp,by="Site") %>%
  mutate(prop_crab=n_crab/total_crab,
         prop_reads=n_reads/total_reads) %>%
  distinct() %>% arrange(n_crab)
```
```{r fig.height=7,fig.width=8}
plotdat <- class_summary_bysite %>%
  dplyr::select(class,Site,n_crab,prop_reads) %>%
  arrange(n_crab) %>%    # First sort by val. This sort the dataframe but NOT the factor levels
  mutate(class=factor(class, levels=unique(class)))
plotdat %>%
  pivot_longer(cols=c(n_crab,prop_reads),names_to="metric",values_to="value") %>%
  mutate(metric=ifelse(metric=="n_crab","Crab Count","Prop Sequence Reads")) %>%
  ggplot(aes(y=value,x=factor(class, levels=unique(plotdat$class)))) +
  facet_grid(cols=vars(metric),scales="free") +
  geom_col(aes(fill=Site), position="dodge") +
  ggtitle("Classes") +
  coord_flip() + theme_bw() + theme(axis.title=element_blank(),
                                    strip.text=element_text(size=12),
                                    axis.text.x=element_text(size=12),
                                    axis.text.y=element_text(size=10))
```
<br>
```{r}
phylum_summary_bysite <- dat.filter %>%
  filter(rank=="species") %>%
  group_by(phylum, Site) %>% 
  summarise(n_crab=length(unique(crab_id)),
            n_reads=sum(nReads)) %>%
  left_join(site.totals.sp,by="Site") %>%
  mutate(prop_crab=n_crab/total_crab,
         prop_reads=n_reads/total_reads) %>%
  distinct() %>% arrange(n_crab)
```
```{r fig.height=5,fig.width=7}
plotdat <- phylum_summary_bysite %>%
  dplyr::select(phylum,Site,n_crab,prop_reads) %>%
  arrange(n_crab) %>%    # First sort by val. This sort the dataframe but NOT the factor levels
  mutate(class=factor(phylum, levels=unique(phylum)))
plotdat %>%
  pivot_longer(cols=c(n_crab,prop_reads),names_to="metric",values_to="value") %>%
  mutate(metric=ifelse(metric=="n_crab","Crab Count","Prop Sequence Reads")) %>%
  ggplot(aes(y=value,x=factor(phylum, levels=unique(plotdat$class)))) +
  facet_grid(cols=vars(metric),scales="free") +
  geom_col(aes(fill=Site), position="dodge") +
  ggtitle("Phyla") +
  coord_flip() + theme_bw() + theme(axis.title=element_blank(),
                                    strip.text=element_text(size=12),
                                    axis.text.x=element_text(size=12),
                                    axis.text.y=element_text(size=10))
```
<br>

What's going on with the "NA" phylum?
```{r}
unique(dat.filter %>% filter(is.na(phylum)) %>% pull(taxon))
```

Some larger algae, some single-celled algae.



